---
title: "R Notebook"
output: html_notebook
---
## Import Libraries
```{r}
library(tidyverse)
library(ggplot2)
library(rpart)
library(rattle) 
library(rpart.plot)
library(RColorBrewer)
library(network)
library(plyr)
library(factoextra)
library(GGally)
library(ggpubr)
library(lubridate)
library(plotly) 
library(htmlwidgets)
library(philentropy)
library(gt)
library(igraph)
library(cvms)
library(tibble)
library(caret)
library(gridExtra)        
```

## Read in Data
```{r}
eotb = read.csv('EOTB_Cleaned.csv')

```

## Clean & Structure Data
```{r}
# Get Rid of non numeric columns
eotb = eotb[,4:11]

# Rename columns
colnames(eotb) = c("Station","Salinity","pH","Dis_Oxygen","Turbidity","Chlorophyll","Temperature","Month")
```

## Change Oxygen Values to labels
```{r}
# Change Oxygen Values to labels - target of study for now. 
ox_lab = eotb

do = ox_lab[,4]

tdo = function(do){
  if(do < 70)
    x = 'Low_Oxygen'
  else if(do >=70 & do < 100)
    x = 'Normal_Oxygen'
  else
    x = "High_Oxygen"
  return(x)
}
dissolved_oxygen = sapply(do, function(x) tdo(x))
ox_lab$Dis_Oxygen = dissolved_oxygen
rm(do,tdo,dissolved_oxygen)

```

## Make char data factors
```{r}
# Change certain data to factors
ox_lab$Station = as.factor(ox_lab$Station)
ox_lab$Month = as.factor(ox_lab$Month)
ox_lab$Dis_Oxygen = as.factor(ox_lab$Dis_Oxygen)

#str(ox_lab)
```

## Dissolved Oxygen Decision Tree 1
```{r}
set.seed(124)
# Get size of dataset
n = nrow(ox_lab)

# Make sizes for training and testing sets
train_set_n = floor(n*(3/4))
test_set_n = n-train_set_n

# Create indicies from sample for training and test sets
train_sample = sample(nrow(ox_lab),train_set_n,replace=FALSE)

# Create training and testing set. 
train_set = ox_lab[train_sample,]

test_set = ox_lab[-train_sample,]

##Remove the labels from test set
test_labs = test_set$Dis_Oxygen
test_set = test_set[,!(names(test_set) %in% c('Dis_Oxygen'))]


##Using information gain
dt_ox1 = rpart(train_set$Dis_Oxygen ~ .,
              data=train_set,
              method = 'class',
              control=rpart.control(minsplit=2, 
                                    minbucket=1) 
                                    #cp=0.015)
              )
#summary(dt_ox1)
```

## Visual for dt_ox1
```{r}
fancyRpartPlot(dt_ox1)
png('dt_ox1.png',width=1000, height=600)
fancyRpartPlot(dt_ox1)
dev.off()
```


## Confusion Matrix Functions
```{r}
### Function below from: https://stackoverflow.com/a/66606295 for confusion matrix visual

conf_matrix <- function(df.true, df.pred, title = "", true.lab ="True Class", pred.lab ="Predicted Class",
                        high.col = 'green3', low.col = 'white') {
  #convert input vector to factors, and ensure they have the same levels
  df.true <- as.factor(df.true)
  df.pred <- factor(df.pred, levels = levels(df.true))
  
  #generate confusion matrix, and confusion matrix as a pecentage of each true class (to be used for color) 
  df.cm <- table(True = df.true, Pred = df.pred)
  df.cm.col <- df.cm / rowSums(df.cm)
  
  #convert confusion matrices to tables, and binding them together
  df.table <- reshape2::melt(df.cm)
  df.table.col <- reshape2::melt(df.cm.col)
  df.table <- left_join(df.table, df.table.col, by =c("True", "Pred"))
  
  #calculate accuracy and class accuracy
  acc.vector <- c(diag(df.cm)) / c(rowSums(df.cm))
  class.acc <- data.frame(Pred = "Class Acc.", True = names(acc.vector), value = acc.vector)
  acc <- sum(diag(df.cm)) / sum(df.cm)
  
  #plot
  ggplot() +
    geom_tile(aes(x=Pred, y=True, fill=value.y),
              data=df.table, size=0.2, color=grey(0.5)) +
    geom_tile(aes(x=Pred, y=True),
              data=df.table[df.table$True==df.table$Pred, ], size=1, color="black", fill = 'transparent') +
    scale_x_discrete(position = "top",  limits = c(levels(df.table$Pred), "Class Acc.")) +
    scale_y_discrete(limits = rev(unique(levels(df.table$Pred)))) +
    labs(x=pred.lab, y=true.lab, fill=NULL,
         title= paste0(title, "\nAccuracy ", round(100*acc, 1), "%")) +
    geom_text(aes(x=Pred, y=True, label=value.x),
              data=df.table, size=4, colour="black") +
    geom_text(data = class.acc, aes(Pred, True, label = paste0(round(100*value), "%"))) +
    scale_fill_gradient(low=low.col, high=high.col, labels = scales::percent,
                        limits = c(0,1), breaks = c(0,0.5,1)) +
    guides(scale = "none") +
    theme_bw() +
    theme(panel.border = element_blank(), legend.position = "bottom",
          axis.text = element_text(color='black'), axis.ticks = element_blank(),
          panel.grid = element_blank(), axis.text.x.top = element_text(angle = 30, vjust = 0, hjust = 0)) +
    coord_fixed()

} 


# Function to create confusion matrix visual & table of stats
run_cm = function(dt,test,test_lab,pos = 1,tit = 'Statistics for Confusion Matrix',subt = NULL){
  # Make DT prediction
  dt_pre = predict(dt,test,type = 'class')
  
  # Create Confusion Matrix
  if(pos == 1){
    cm = confusionMatrix(dt_pre,test_lab, positive = 'true')
    
    # Convert statistics to DF & pull out desired statistics
    stats = as.data.frame(t(as.matrix(cm$byClass)))
    stats = stats[c(1,2,5,7),]}
  
  
  else{
    cm = confusionMatrix(dt_pre,test_lab)
    
    # Convert statistics to DF & pull out desired statistics
    stats = as.data.frame(cm$byClass[c(1,2,5,7)])
    colnames(stats) = ' '}
    
    
  
  # Create table with gt package
  stats_tb = gt(stats,rownames_to_stub = T)
  stats_tb = stats_tb %>%
  
  # Create title & Subtitle
    tab_header(
      title = tit,
      subtitle = subt)
       
  

  # Save table as variable
  cm_stats = stats_tb
  
  # Create confusion matrix visual and save as variable
  cm_viz = conf_matrix(test_lab,dt_pre)
  
  # Return CM visual & stats as a list
  return(list(cm_viz,cm_stats))
}



```


## Confusion Matrix for dt_ox1
```{r}
# Get confusion matrix visual & stats for dt_ox1
(dt_ox1_viz = run_cm(dt_ox1,
                     test_set,
                     test_labs,
                     tit = 'Confusion Matrix Statistics by Class',
                     subt = "For Predicting Dissolved Oxygen Levels"))

# save plot & table
ggsave('dt_ox1_cm.png',plot = dt_ox1_viz[[1]])
gtsave(dt_ox1_viz[[2]],'dt_ox1_cm_stats.html')


```

## Decision Tree for oxygen 2 (subset of data)
```{r}
# Subset oxygen data
ox_lab2 = ox_lab[,!(names(ox_lab) %in% c('Station','pH'))]

set.seed(124)

# Get size of dataset
n = nrow(ox_lab2)

# Make sizes for training and testing sets
train_set_n = floor(n*(3/4))
test_set_n = n-train_set_n

# Create indicies from sample for training and test sets
train_sample = sample(nrow(ox_lab2),train_set_n,replace=FALSE)

# Create training and testing set. 
train_set = ox_lab2[train_sample,]

test_set = ox_lab2[-train_sample,]

##Remove the labels from test set
test_labs = test_set$Dis_Oxygen
test_set = test_set[,!(names(test_set) %in% c('Dis_Oxygen'))]


##Using information gain
dt_ox2 = rpart(train_set$Dis_Oxygen ~ .,
              data=train_set,
              method = 'class',
              control=rpart.control(minsplit=2, 
                                    minbucket=1, 
                                    cp=0.00725)
              )
#summary(dt_ox2)

```

## Visual for dt_ox2
```{r}

fancyRpartPlot(dt_ox2)
png('dt_ox2.png',width=1000, height=600)
fancyRpartPlot(dt_ox2)
dev.off()

```

## Confusion Matrix & stats for dt_ox2
```{r}
# Get confusion matrix visual & stats for dt_ox1
(dt_ox2_viz = run_cm(dt_ox2,
                     test_set,
                     test_labs,
                     tit = 'Confusion Matrix Statistics by Class',
                     subt = "For Predicting Dissolved Oxygen Levels"))

# save plot & table
ggsave('dt_ox2_cm.png',plot = dt_ox2_viz[[1]])
gtsave(dt_ox2_viz[[2]],'dt_ox2_cm_stats.html')
```

## Next target of interest is pH
```{r}
ph_lab = eotb

ph = round(ph_lab[3],3)
ph = ph[ph>=0 & ph <= 14]

# View pH range
#summary(ph)

tph = function(ph){
  if(ph < 7)
    x  = 'Acidic'
#  else if(ph >= 6.75 & ph <= 7.25)
#    x = 'Neutral'
  else if(ph >= 7)
    x = 'Not Acidic'
  return(x)
}
ph = sapply(ph, function(x) tph(x))
ph_lab$pH = ph
rm(ph,tph)
#table(ph)

# Check structure of DF
#str(ph_lab)

# Change necessary data types to factors
ph_lab$Station = as.factor(ph_lab$Station)
ph_lab$pH = as.factor(ph_lab$pH)
ph_lab$Month = as.factor(ph_lab$Month)

#table(ph_lab$pH)
```

## pH Decision Tree
```{r}
set.seed(124)

# Get size of dataset
n = nrow(ph_lab)

# Make sizes for training and testing sets
train_set_n = floor(n*(3/4))
test_set_n = n-train_set_n

# Create indicies from sample for training and test sets
train_sample = sample(nrow(ph_lab),train_set_n,replace=FALSE)

# Create training and testing set. 
train_set = ph_lab[train_sample,]

test_set = ph_lab[-train_sample,]

##Remove the labels from test set
test_labs = test_set$pH
test_set = test_set[,!(names(test_set) %in% c('pH'))]


##Using information gain
dt_ph = rpart(train_set$pH ~ .,
              data=train_set,
              method = 'class',
              control=rpart.control(minsplit=2, 
                                    minbucket=1) 
                                    #cp=0.00725)
              )
summary(dt_ph)
```

## Visual for dt_ph
```{r}
fancyRpartPlot(dt_ph)
png('dt_ph.png',width=1000, height=600)
fancyRpartPlot(dt_ph)
dev.off()

```

## Confusion Matrix dt_ph
```{r}
# Get confusion matrix visual & stats for dt_ox1
(dt_ph_viz = run_cm(dt_ph,
                    test_set,
                    test_labs,
                    pos = 0,
                    subt = "For Predicting pH"))

  
  # Create Confusion Matrix
# save plot & table
ggsave('dt_ph_cm.png',plot = dt_ph_viz[[1]])
gtsave(dt_ph_viz[[2]],'dt_ph_cm_stats.html')
```

## DT for pH independet of Station Variable
```{r}
ph_lab2 = ph_lab[,!(names(ph_lab) %in% c('Station'))]

set.seed(124)

# Get size of dataset
n = nrow(ph_lab2)

# Make sizes for training and testing sets
train_set_n = floor(n*(3/4))
test_set_n = n-train_set_n

# Create indicies from sample for training and test sets
train_sample = sample(nrow(ph_lab2),train_set_n,replace=FALSE)

# Create training and testing set. 
train_set = ph_lab2[train_sample,]

test_set = ph_lab2[-train_sample,]

##Remove the labels from test set
test_labs = test_set$pH
test_set = test_set[,!(names(test_set) %in% c('pH'))]


##Using information gain
dt_ph2 = rpart(train_set$pH ~ .,
              data=train_set,
              method = 'class',
              control=rpart.control(minsplit=2, 
                                    minbucket=1) 
                                    #cp=0.00725)
              )
summary(dt_ph2)
```

## Visual for dt_ph2 (independent of station variable)
```{r}
fancyRpartPlot(dt_ph2)
png('dt_ph2.png',width=1000, height=600)
fancyRpartPlot(dt_ph2)
dev.off()
  
```

## Confusion Matrix dt_ph2 (independent of station variable)
```{r}
# Get confusion matrix visual & stats for dt_ox1
(dt_ph_viz2 = run_cm(dt_ph2,
                     test_set,
                     test_labs,
                     pos = 0,
                     subt = "For Predicting pH"))

# save plot & table
ggsave('dt_ph_cm2.png',plot = dt_ph_viz2[[1]])
gtsave(dt_ph_viz2[[2]],'dt_ph_cm_stats2.html')
```

## Next target of interest is chlorophyll
```{r}
cl_lab = eotb

cl = cl_lab[6]
#summary(cl)
cl = cl[cl > 0 & cl < 500]

tcl = function(cl){
  if(cl < 15)
    x  = 'Not Harmful CL'

  else if(cl >= 15)
    x = 'Harmul CL'
  return(x)
}
cl = sapply(cl, function(x) tcl(x))
cl_lab$Chlorophyll = cl
rm(cl,tcl)


# Check structure of DF
#str(cl_lab)

# Change necessary data types to factors
cl_lab$Station = as.factor(cl_lab$Station)
cl_lab$Chlorophyll = as.factor(cl_lab$Chlorophyll)
cl_lab$Month = as.factor(cl_lab$Month)

table(cl_lab$Chlorophyll)
```

## Chlorophyll Decision Tree
```{r}
set.seed(120)

# Get size of dataset
n = nrow(cl_lab)

# Make sizes for training and testing sets
train_set_n = floor(n*(3/4))
test_set_n = n-train_set_n

# Create indicies from sample for training and test sets
train_sample = sample(nrow(cl_lab),train_set_n,replace=FALSE)

# Create training and testing set. 
train_set = cl_lab[train_sample,]

test_set = cl_lab[-train_sample,]

##Remove the labels from test set
test_labs = test_set$Chlorophyll
test_set = test_set[,!(names(test_set) %in% c('Chlorophyll'))]


##Using information gain
dt_cl = rpart(train_set$Chlorophyll ~ .,
              data=train_set,
              method = 'class',
              control=rpart.control(minsplit=2, 
                                    minbucket=1,
                                    cp = .0125)
                                    
              )
#summary(dt_cl)
```

## Visual for dt_cl
```{r}
fancyRpartPlot(dt_cl)
png('dt_cl.png',width=1000, height=600)
fancyRpartPlot(dt_cl)
dev.off()

```

## Confusion Matrix for dt_cl
```{r}
# Get confusion matrix visual & stats for dt_ox1
(dt_cl_viz = run_cm(dt_cl,
                    test_set,
                    test_labs,
                    pos = 0,
                    subt = "For Predicting Chlorophyll"))

  
  # Create Confusion Matrix
# save plot & table
ggsave('dt_cl_cm.png',plot = dt_cl_viz[[1]])
gtsave(dt_cl_viz[[2]],'dt_cl_cm_stats.html')
```

## Chlorophyll Decision Tree Independent of Station 
```{r}
cl_lab2 = cl_lab[,!(names(cl_lab) %in% c('Station'))]

set.seed(120)

# Get size of dataset
n = nrow(cl_lab2)

# Make sizes for training and testing sets
train_set_n = floor(n*(3/4))
test_set_n = n-train_set_n

# Create indicies from sample for training and test sets
train_sample = sample(nrow(cl_lab2),train_set_n,replace=FALSE)

# Create training and testing set. 
train_set = cl_lab2[train_sample,]

test_set = cl_lab2[-train_sample,]

##Remove the labels from test set
test_labs = test_set$Chlorophyll
test_set = test_set[,!(names(test_set) %in% c('Chlorophyll'))]


##Using information gain
dt_cl2 = rpart(train_set$Chlorophyll ~ .,
              data=train_set,
              method = 'class',
              control=rpart.control(minsplit=2, 
                                    minbucket=1,
                                    cp = .0205)
                                    
              )
summary(dt_cl2)
```

## Visual for dt_cl
```{r}
fancyRpartPlot(dt_cl2)
png('dt_cl2.png',width=1000, height=600)
fancyRpartPlot(dt_cl2)
dev.off()

```

## Confusion Matrix for dt_cl
```{r}
# Get confusion matrix visual & stats for dt_ox1
(dt_cl_viz2 = run_cm(dt_cl2,
                    test_set,
                    test_labs,
                    pos = 0,
                    subt = "For Predicting Chlorophyll"))

  
  # Create Confusion Matrix
# save plot & table
ggsave('dt_cl_cm2.png',plot = dt_cl_viz2[[1]])
gtsave(dt_cl_viz2[[2]],'dt_cl_cm_stats2.html')
```