---
title: "R Notebook"
output: html_notebook
---

###########################
Notes about output:

Upon output - there will be a single data set for each season spanning from the 
09-10 season to the 20-21 season. Each dataset contains the player, his team, his 
salary, and the number of games he played that season. My thoughts for using this 
for analysis would be to label teams based on win percentage as a good or bad team
and then we could do some hypothesis testing on whether the good teams have a higher
or lower median salary, top salary, Q3 salary etc. 

###########################

### Data URLs ###
https://hoopshype.com/salaries/players/2020-2021/
https://www.nba.com/stats/players/traditional/?sort=PTS&dir=-1&Season=2020-21&SeasonType=Regular%20Season

#################

## Libraries
```{r}
library(ggplot2)
library(XLConnect)
library(tidyverse)
library(ggimage)
library(jsonlite)
library(httr)
library(rvest)
library(gghighlight)
`%!in%` <- Negate(`%in%`)
options(java.parameters = "- Xmx1024m")
```


## Importing Datasets
```{r}
l_sals = c()
years = c('09_10',"10_11","11_12","12_13","13_14","14_15","15_16","16_17","17_18","18_19","19_20","20_21")
for(i in 1:12){
  sals = readWorksheetFromFile("raw_salary_data.xlsx",sheet = i,header = FALSE)
  d_var_1 = paste0('sals_',years[i])
  sals = sals[,2:3]
  colnames(sals) = c('player','salary')
  assign(d_var_1, sals)
  xlcFreeMemory()

  
}
l_stats = c()
for(i in 1:12){
  stats = readWorksheetFromFile("raw_player+team_data.xlsx",sheet = i)
  d_var_2 = paste0('stats_',years[i])
  stats = stats[,c(2:3,5)]
  assign(d_var_2, stats)
  xlcFreeMemory()
}

```

## Join Datasets
```{r}
sal_join = function(sal,stat){
  sal = sal %>%
    left_join(stat, by = c('player' = 'PLAYER')) %>%
    drop_na() %>%
    arrange(TEAM)
  return(sal)
}
sals_09_10 = sal_join(sals_09_10,stats_09_10)
sals_10_11 = sal_join(sals_10_11,stats_10_11)
sals_11_12 = sal_join(sals_11_12,stats_11_12)
sals_12_13 = sal_join(sals_12_13,stats_12_13)
sals_13_14 = sal_join(sals_13_14,stats_13_14)
sals_14_15 = sal_join(sals_14_15,stats_14_15)
sals_15_16 = sal_join(sals_15_16,stats_15_16)
sals_16_17 = sal_join(sals_16_17,stats_16_17)
sals_17_18 = sal_join(sals_17_18,stats_17_18)
sals_18_19 = sal_join(sals_18_19,stats_18_19)
sals_19_20 = sal_join(sals_19_20,stats_19_20)
sals_20_21 = sal_join(sals_20_21,stats_20_21)
sals = list(sals_09_10,sals_10_11,sals_11_12,sals_12_13,sals_13_14,
            sals_14_15,sals_15_16,sals_16_17,sals_17_18,sals_18_19,
            sals_19_20,sals_20_21)




```

## Write to csvs
```{r}
for(i in 1:12){
  fn = paste0('salaries_',years[i],'.csv')
  write_csv(sals[[i]],fn)
}

```


```{r}
df = data.frame(player=character(),
                salary=numeric(), 
                TEAM=character(),
                GP=numeric(),
                year=character(),
                stringsAsFactors=FALSE) 

for(i in 1:12){
  x = sals[[i]]
  x$year = years[i]

  df = rbind(df,x)
  
}
write_csv(df,'all_seasons_salaries.csv')

```



