---
title: "R Notebook"
output: html_notebook
---

## Import Libraries
```{r}

library(tidyverse)
library(ggplot2)
library(rpart)
library(rattle) 
library(rpart.plot)
library(RColorBrewer)
library(network)
library(plyr)
library(factoextra)
library(GGally)
library(ggpubr)
library(lubridate)
library(plotly) 
library(htmlwidgets)
library(philentropy)
library(gt)
library(igraph)
library(cvms)
library(tibble)
library(caret)
library(gridExtra)     
library(caretEnsemble)
library(tm)
library(slam)
library(quanteda)
library(SnowballC)
library(arules)
library(proxy)
library(cluster)
library(Matrix)
library(plyr) 
library(mclust) 
library(naivebayes)
library(psych)
library(mice)
library(e1071)
library(mlr)
```


## Confusion Matrix Function
```{r}
### Function below from: https://stackoverflow.com/a/66606295 for confusion matrix visual

conf_matrix <- function(df.true, df.pred, true.lab ="True Class", pred.lab ="Predicted Class",
                        high.col = 'green3', low.col = 'white',title = "Confusion Matrix") {
  #convert input vector to factors, and ensure they have the same levels
  df.true <- as.factor(df.true)
  df.pred <- factor(df.pred, levels = levels(df.true))
  
  #generate confusion matrix, and confusion matrix as a pecentage of each true class (to be used for color) 
  df.cm <- table(True = df.true, Pred = df.pred)
  df.cm.col <- df.cm / rowSums(df.cm)
  
  #convert confusion matrices to tables, and binding them together
  df.table <- reshape2::melt(df.cm)
  df.table.col <- reshape2::melt(df.cm.col)
  df.table <- left_join(df.table, df.table.col, by =c("True", "Pred"))
  
  #calculate accuracy and class accuracy
  acc.vector <- c(diag(df.cm)) / c(rowSums(df.cm))
  class.acc <- data.frame(Pred = "Class Acc.", True = names(acc.vector), value = acc.vector)
  acc <- sum(diag(df.cm)) / sum(df.cm)
  
  #plot
  ggplot() +
    geom_tile(aes(x=Pred, y=True, fill=value.y),
              data=df.table, size=0.2, color=grey(0.5)) +
    geom_tile(aes(x=Pred, y=True),
              data=df.table[df.table$True==df.table$Pred, ], size=1, color="black", fill = 'transparent') +
    scale_x_discrete(position = "top",  limits = c(levels(df.table$Pred), "Class Acc.")) +
    scale_y_discrete(limits = rev(unique(levels(df.table$Pred)))) +
    labs(x=pred.lab, 
         y=true.lab, 
         fill=NULL,
         title = title,
         caption = paste0("Total Accuracy: ", round(100*acc, 1), "%")) +
    geom_text(aes(x=Pred, y=True, label=value.x),
              data=df.table, size=4, colour="black") +
    geom_text(data = class.acc, aes(Pred, True, label = paste0(round(100*value), "%"))) +
    scale_fill_gradient(low=low.col, high=high.col, labels = scales::percent,
                        limits = c(0,1), breaks = c(0,0.5,1)) +
    guides(scale = "none") +
    theme_bw() +
    theme(panel.border = element_blank(), legend.position = "bottom",
          axis.text = element_text(color='black'), axis.ticks = element_blank(),
          panel.grid = element_blank(), axis.text.x.top = element_text(angle = 30, vjust = 0, hjust = 0),
          plot.title = element_text(size=17, face="bold", 
      margin = margin(10, 0, 10, 0),hjust =.5),
      plot.subtitle = element_text(hjust =.5)) +
    coord_fixed()

} 

```

## Read in Data
```{r}
eotb = read.csv('EOTB_Cleaned.csv')

```

## Clean & Structure Data
```{r}
# Get Rid of non numeric/factor columns
eotb = eotb[,5:11]

# Rename columns
colnames(eotb) = c("Salinity","pH","Dis_Oxygen","Turbidity","Chlorophyll","Temperature","Month")
#eotb$Station = as.factor(eotb$Station)
eotb$Month = as.factor(eotb$Month)
```



## Change Oxygen Values to labels
```{r}
# Change Oxygen Values to labels - target of study for now. 
ox_lab = eotb

do = ox_lab[,3]

tdo = function(do){
  if(do < 70)
    x = 'Low_Oxygen'
  else if(do >=70 & do < 100)
    x = 'Normal_Oxygen'
  else
    x = "High_Oxygen"
  return(x)
}
dissolved_oxygen = sapply(do, function(x) tdo(x))
ox_lab$Dis_Oxygen = as.factor(dissolved_oxygen)
rm(do,tdo,dissolved_oxygen)

```


## View amounts of each label
```{r}
#table(ox_lab$Dis_Oxygen)

```

## Scale the data
```{r}
ox_lab$Salinity = scale(ox_lab$Salinity)
ox_lab$pH = scale(ox_lab$pH)
ox_lab$Turbidity = scale(ox_lab$Turbidity)
ox_lab$Chlorophyll = scale(ox_lab$Chlorophyll)
ox_lab$Temperature = scale(ox_lab$Temperature)

```

## Split into test & training set
```{r}
set.seed = 123
# Need even amounts in training set to train data:, low_oxygen has fewest labels so least common denominator

# Create column of index rows
ox_lab$row_num = 1:nrow(ox_lab)

# Find amount of low oxygen values
l = nrow(ox_lab[ox_lab$Dis_Oxygen == 'Low_Oxygen',])

# Make training size 70% of set
train_size = l*.7

# Extract indicies of even amount of each label for training
high_o_sam = sample(ox_lab$row_num[ox_lab$Dis_Oxygen == 'High_Oxygen'],train_size)
norm_o_sam = sample(ox_lab$row_num[ox_lab$Dis_Oxygen == 'Normal_Oxygen'],train_size)
low_o_sam = sample(ox_lab$row_num[ox_lab$Dis_Oxygen == 'Low_Oxygen'],train_size)

# Combine Indicies
train_ind = c(high_o_sam,norm_o_sam,low_o_sam)

# Remove row indicies from df
ox_lab = ox_lab[,1:7]

# Get training and Testing Set
train_set = ox_lab[train_ind,]
test_set = ox_lab[-train_ind,]

# Save Labels in variables
train_labs = train_set$Dis_Oxygen
test_labs = test_set$Dis_Oxygen

# Remove Labels from sets
train_set = train_set[,c(1:2,4:7)]
test_set = test_set[,c(1:2,4:7)]
rm(ox_lab,high_o_sam,norm_o_sam,low_o_sam,l,train_size,train_ind)
```

## Run Naive Bayes on Test Set
```{r}
model = e1071::naiveBayes(train_set, 
                       train_labs, 
                       laplace = 1)


model_predict = predict(model, test_set)


```

```{r}
#model
cm = conf_matrix(test_labs,model_predict)
ggsave('ox_cm.png',plot = cm)

```

## Create Tables to show relevant statistics
```{r}
# Create variables to fill with low oxygen level info
VARIABLE = c('Salinity','pH','Turbidity','Chlorophyll','Temperature','Month')
MEAN = c()
SD = c()

# Fill in variables with low ox level info
for (i in 1:5){
  tb = model$tables[[VARIABLE[i]]]
  MEAN = append(MEAN,round(tb[2,1],3))
  SD = append(SD,round(tb[2,2],3))
}

# Create variable with month cond. probs for low ox levels
MONTH = model$tables[[VARIABLE[6]]]
MONTH = round(MONTH[2,],5)

# create df and change col names for table
df = data.frame(VARIABLE[1:5],MEAN,SD)
colnames(df) = c('Variable','Mean','Standard Deviation')

# Create table for mean and SD of variables
stats_tb = gt(df,rownames_to_stub = T)
stats_tb = stats_tb %>%
  
# Create title & Subtitle
  tab_header(
    title = 'Mean & Standard Deviation of Predicted Variables for Low Oxygen Levels',
    subtitle = 'All Variables are Normalized')%>%
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = `Standard Deviation`,
      rows = `Standard Deviation` < 0.8
    ))
       
  

# Save table as variable
ox_stats = stats_tb
#ox_stats
gtsave(ox_stats,'Ox_stats_tab.html')


# Create DF for cond. probs by month, change col names for table
df = data.frame(c('January','February','March','April','May','June','July','August','September','October','November','December'),MONTH)
colnames(df) = c('Month','Conditional Probability')

# Create Table
mon_tb= gt(df)
mon_tb = mon_tb %>%
  tab_header(
    title = 'Conditional Probability of Predicted Low Oxygen Levels by Month') %>%
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = `Conditional Probability`,
      rows = `Conditional Probability` > 10^(-1)
    ))
ox_mon_tb = mon_tb
#ox_mon_tb
gtsave(ox_mon_tb,'Ox_mon_tab.html')
  

rm(VARIABLE,MEAN,SD,tb,MONTH,df,stats_tb,ox_stats,mon_tb,ox_mon_tb,i)

```



## Change Chlorophyll values
```{r}
cl_lab = eotb

cl = cl_lab[5]
#summary(cl)
cl = cl[cl > 0 & cl < 500]

tcl = function(cl){
  if(cl < 15)
    x  = 'Not Harmful CL'

  else if(cl >= 15)
    x = 'Harmful CL'
  return(x)
}
cl = sapply(cl, function(x) tcl(x))
cl_lab$Chlorophyll = as.factor(cl)
rm(cl,tcl)

# Change necessary data types to factors
cl_lab$Month = as.factor(cl_lab$Month)
```


## View amounts of each label
```{r}
#table(cl_lab$Chlorophyll)

```

## Scale the data
```{r}
cl_lab$Salinity = scale(cl_lab$Salinity)
cl_lab$pH = scale(cl_lab$pH)
cl_lab$Turbidity = scale(cl_lab$Turbidity)
cl_lab$Dis_Oxygen = scale(cl_lab$Dis_Oxygen)
cl_lab$Temperature = scale(cl_lab$Temperature)

```

## Split into test & training set
```{r}
set.seed = 123
# Need even amounts in training set to train data:, harmful CL has fewest labels so least common denominator

# Create column of index rows
cl_lab$row_num = 1:nrow(cl_lab)

# Find amount of harmful CL values
l = nrow(cl_lab[cl_lab$Chlorophyll == 'Harmful CL',])

# Make training size 70% of set
train_size = l*.7

# Extract indicies of even amount of each label for training
harm = sample(cl_lab$row_num[cl_lab$Chlorophyll == 'Harmful CL'],train_size)
not_harm = sample(cl_lab$row_num[cl_lab$Chlorophyll == 'Not Harmful CL'],train_size)

# Combine Indicies
train_ind = c(harm,not_harm)

# Remove row indicies from df
cl_lab = cl_lab[,1:7]

# Get training and Testing Set
train_set = cl_lab[train_ind,]
test_set = cl_lab[-train_ind,]

# Save Labels in variables
train_labs = train_set$Chlorophyll
test_labs = test_set$Chlorophyll

# Remove Labels from sets
train_set = train_set[,c(1:4,6:7)]
test_set = test_set[,c(1:4,6:7)]

rm(harm,not_harm,cl_lab,l,train_ind,train_size)
```

## Run Naive Bayes on Test Set
```{r}
model = e1071::naiveBayes(train_set, 
                       train_labs, 
                       laplace = 1)


model_predict = predict(model, test_set)


```

```{r}
#model
cm = conf_matrix(test_labs,model_predict)
#cm
ggsave('cl_cm.png',plot = cm)
rm(cm)
```

## Create Tables to show relevant statistics
```{r}
# Create variables to fill with relevant table values
VARIABLE = c('Salinity','pH','Dis_Oxygen','Turbidity','Temperature','Month')
MEAN = c()
SD = c()

# Fill harmful chlorophyll values into created variables
for (i in 1:5){
  tb = model$tables[[VARIABLE[i]]]
  MEAN = append(MEAN,round(tb[1,1],3))
  SD = append(SD,round(tb[1,2],3))
}

# Put values of harmful chlorophyll cond. probs into variable
MONTH = model$tables[[VARIABLE[6]]]
MONTH = round(MONTH[1,],5)

# Create dataframe with variables, means and standard deviations
df = data.frame(VARIABLE[1:5],MEAN,SD)

# Change Column names for table
colnames(df) = c('Variable','Mean','Standard Deviation')

# Create table
stats_tb = gt(df,rownames_to_stub = T)
stats_tb = stats_tb %>%
  
# Create title & Subtitle
  tab_header(
    title = 'Mean & Standard Deviation of Predicted Variables for Harmful Chlorophyll Levels',
    subtitle = 'All Variables are Normalized')%>%
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = `Standard Deviation`,
      rows = `Standard Deviation` < 0.8
    ))
       
  

# Save table as variable
cl_stats = stats_tb
#cl_stats
 


# Create separate data frame for months and cond. probs
df = data.frame(c('January','February','March','April','May','June','July','August','September','October','November','December'),MONTH)

# Change col names for table
colnames(df) = c('Month','Conditional Probability')

# Create Table
mon_tb= gt(df)
mon_tb = mon_tb %>%
  tab_header(
    title = 'Conditional Probability of Predicted Low Oxygen Levels by Month') %>%
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = `Conditional Probability`,
      rows = `Conditional Probability` > .15
    ))
cl_mon_tb = mon_tb
#cl_mon_tb
gtsave(cl_mon_tb,'CL_mon_tab.html')
  


rm(VARIABLE,MEAN,SD,tb,MONTH,df,stats_tb,cl_stats,mon_tb,cl_mon_tb,i)

```


