---
title: "R Notebook"
output: html_notebook
---


## Import Libraries
```{r}
library(tidyverse)
library(plyr)
library(ggplot2)
library(factoextra)
library(GGally)
library(ggpubr)
library(lubridate)
library(cluster)
library(fpc)
library(dbscan)
library(plotly) 
library(ggdendro)
library(htmlwidgets)
library(philentropy)
library(arules)
library(arulesViz)
library(gt)
library(igraph)

```

## Read in data 
```{r}

eotb = read.csv('EOTB_Cleaned.csv')

# Remove labels and dates
eotb = eotb[4:11]
```


## Create Table with category specifications
```{r}
Salinity = c('Very Low (<= .25)',
             "Low (>.25 , <=1)",
             "Normal (>1 , <=10)",
             "High (>10 , <=20)",
             "Very High (>20)")
pH = c('Acidic (<6)',
       "Slightly Acidic (>=6 , <=6.75)",
       "Neutral (>6.75 , <7.25)",
       "Slightly Basic (>=7.25 , <=8)",
       "Basic (>8)")
Dis_O = c('Very Low (<50)',
          "Low (>=50 , <75)",
          "Normal (>=75 , <95)",
          "High (>=95 , <125)",
          "Abundant (>=125)")
Turbidity = c('Very Low (<1)',
              "Low (>=1 , < 3)",
              "Normal (>=3 , <7)",
              "High (>=7 , <15)",
              "Very High (>=15)")
Chlorophyll = c('Very Low (<1)',
              "Low (>=1 , < 4)",
              "Normal (>=4 , <11)",
              "High (>=11 , <20)",
              "Very High (>=20)")
Water = c('Very Cold (<40)',
              "Cold (>=40 , < 60)",
              "Temperate (>=60 , <70)",
              "Warm (>=70 , <80)",
              "Hot (>=80)")
specs_df = data.frame(Salinity,pH,Dis_O,Turbidity,Chlorophyll,Water)
colnames(specs_df) = c("Salinity","pH","Dissolved Oxygen","Turbidity","Chlorophyll","Water")



# Create table with gt package
specs = gt(specs_df)
specs = specs %>%
  
  # Create title & Subtitle
  tab_header(
    title = 'Data Binning Specifications'
  ) %>%
  
  # Align Columns
  cols_align(
  align = "center",#c("center", "center", "center", "center", "center", "center"),
  columns = everything()
  )

specs
gtsave(specs,'specs_table.html')

# Remove unneeded variables from environment to keep it clear
rm(list=(c("Salinity","pH","Dis_O","Turbidity","Chlorophyll","Water","specs","specs_df")))
```



## Transform Record data to transaction data
```{r}
# Station already in correct format. 
# Put column in new dataframe
tran = data.frame(eotb[,1])
```

# Salinity
```{r}
sal = eotb[,2]

# View Salinity range
#summary(sal)
#summary(sal[sal <10])

# It is hard to gauge a normal range but by doing some research, these are approximately ranges for measures of salinity. 
tsal = function(sal){
  if(sal <= .25)
    x  = 'Very_Low_Salinity'
  else if(sal > .25 & sal <= 1)
    x = 'Low_Salinity'
  else if(sal > 1 & sal <= 10)
    x = 'Normal_Salinity'
  else if(sal > 10 & sal <= 20)
    x = 'High_Salinity'
  else if(sal > 20)
    x = 'Very_High_Salinity'
  return(x)
}
sal = sapply(sal, function(x) tsal(x))
tran = cbind(tran,sal)
rm(sal,tsal)
#table(tran$sal)
```
## pH
```{r}
ph = eotb[,3]

# View pH range
#summary(ph)


tph = function(ph){
  if(ph < 6)
    x  = 'Acidic'
  else if(ph >= 6 & ph <= 6.75)
    x = 'Slightly_Acidic'
  else if(ph > 6.75 & ph < 7.25)
    x = 'Neutral'
  else if(ph >= 7.25 & ph <= 8)
    x = 'Slightly_Basic'
  else if(ph > 8)
    x = 'Basic'
  return(x)
}
ph = sapply(ph, function(x) tph(x))
tran = cbind(tran,ph)
rm(ph,tph)
#table(tran$ph)
```
#Dissolved Oxygen
```{r}
do = eotb[,4]

# View pH range
#summary(do)


tdo = function(do){
  if(do < 50)
    x  = 'Very_Low_Oxygen'
  else if(do >= 50 & do < 75)
    x = 'Low_Oxygen'
  else if(do >= 75 & do < 95)
    x = 'Normal_Oxygen'
  else if(do >= 95 & do < 125)
    x = 'High_Oxygen'
  else if(do >= 125)
    x = 'Abundant_Oxygen'
  return(x)
}
dissolved_oxygen = sapply(do, function(x) tdo(x))
tran = cbind(tran,dissolved_oxygen)
rm(do,tdo,dissolved_oxygen)
#table(tran$dissolved_oxygen)
```

#Turbidity
```{r}
tur = eotb[,5]

# View pH range
#summary(tur)

ttur = function(tur){
  if(tur < 1)
    x  = 'Very_Low_Turbidity'
  else if(tur >= 1 & tur < 3)
    x = 'Low_Turbidity'
  else if(tur >= 3 & tur < 7)
    x = 'Normal_Turbidity'
  else if(tur >= 7 & tur < 15)
    x = 'High_Turbidity'
  else if(tur >= 15)
    x = 'Very_High_Turbidity'
  return(x)
}
turbidity = sapply(tur, function(x) ttur(x))
tran = cbind(tran,turbidity)
rm(tur,ttur,turbidity)
#table(tran$turbidity)
```

#Chlorophyll
```{r}
chl = eotb[,6]

# View pH range
#summary(chl)

tchl = function(chl){
  if(chl < 1)
    x  = 'Very_Low_Chlorophyll'
  else if(chl >= 1 & chl < 4)
    x = 'Low_Chlorophyll'
  else if(chl >= 4 & chl < 11)
    x = 'Normal_Chlorophyll'
  else if(chl >= 11 & chl < 20)
    x = 'High_Chlorophyll'
  else if(chl >= 20)
    x = 'Very_High_Chlorophyll'
  return(x)
}
chlorophyll = sapply(chl, function(x) tchl(x))
tran = cbind(tran,chlorophyll)
#table(tran$chlorophyll)
rm(chl,tchl,chlorophyll)
```

#Water Temperature
```{r}
temp = eotb[,7]

# View pH range
#summary(temp)


ttemp = function(temp){
  if(temp < 40)
    x  = 'Very_Cold'
  else if(temp >= 40 & temp < 60)
    x = 'Cold'
  else if(temp >= 60 & temp < 70)
    x = 'Temperate'
  else if(temp >= 70 & temp < 80)
    x = 'Warm'
  else if(temp >= 80)
    x = 'Hot'
  return(x)
}
water_temp = sapply(temp, function(x) ttemp(x))
tran = cbind(tran,water_temp)
rm(temp,ttemp,water_temp)
#table(tran$water_temp)
```

# Months
```{r}
mon = eotb[,8]


tmon = function(mon){
  if(mon == 1)
    x  = 'January'
  else if(mon == 2)
    x = 'February'
  else if(mon == 3)
    x = 'March'
  else if(mon == 4)
    x = 'April'
  else if(mon == 5)
    x = 'May'
  else if(mon == 6)
    x = 'June'
  else if(mon == 7)
    x = 'July'
  else if(mon == 8)
    x = 'August'
  else if(mon == 9)
    x = 'September'
  else if(mon == 10)
    x = 'October'
  else if(mon == 11)
    x = 'November'
  else if(mon == 12)
    x = 'December'
  

  return(x)
}

Month = sapply(mon, function(x) tmon(x))
tran = cbind(tran,Month)
rm(mon,tmon,Month)
#table(tran$Month)

colnames(tran) = c('Station','Salinity','pH','Dis_Oxygen','Turbidity','Chlorophyll','Water_Temp','Month')

```

## Write to csv file
```{r}

write.csv(tran,'eotb_grouped.csv')
```

## Create Transaction Data  
```{r}
eotb_trans = read.transactions("eotb_grouped.csv",
                               sep =",", 
                               format("basket"),
                               rm.duplicates = TRUE)


```

## Item Frequency Plot
```{r}

png(file="freq_plot.png",width = 1000,height = 600)
freq_plot = itemFrequencyPlot(eotb_trans, 
                              topN=15,  
                              cex.names=1.5, 
                              cex.axis = 1.5,
                              cex.lab = 1.5,
                              main = "Item Relative Frequency Plot", 
                              col = "navy")
dev.off()
rm(freq_plot)
```

## Create Association Rules
```{r}
#inspect(rows[1:10])
eotb_rules = arules::apriori(eotb_trans, 
        parameter = list(support=.075, conf=.6, maxlen=6))

inspect(eotb_rules[1:20])
##  Sort by Conf
eotb_rules_conf = sort(eotb_rules, by="confidence", decreasing=TRUE)
inspect(eotb_rules_conf[1:20])
## Sort by Sup
eotb_rules_sup = sort(eotb_rules, by="support", decreasing=TRUE)
inspect(eotb_rules_sup[1:20])
## Sort by Lift
eotb_rules_lift = sort(eotb_rules, by="lift", decreasing=TRUE)
inspect(eotb_rules_lift[1:20])
```

## Creating Table of Confidence Values
```{r}
# Create dataframe with rules sorted by confidence
eotb_conf = DATAFRAME(eotb_rules_conf, separate = TRUE)

# Make rules more visually appealing
eotb_conf[] = lapply(eotb_conf, gsub, pattern='[{]', replacement='')
eotb_conf[] = lapply(eotb_conf, gsub, pattern='[}]', replacement='')
eotb_conf[] = lapply(eotb_conf, gsub, pattern='[_]', replacement=' ')
eotb_conf[] = lapply(eotb_conf, gsub, pattern='[,]', replacement=' & ')

# Get columns for table & change column names
eotb_conf = eotb_conf[,c('LHS','RHS','confidence','support','lift')]
colnames(eotb_conf) = c('Source','Target', 'Confidence','Support','Lift')

# Round numbers for table
eotb_conf[,3:5] = as.numeric(unlist(eotb_conf[,3:5]))
eotb_conf[,3:5] = round(eotb_conf[,3:5],3)

# Get top 15 rules
disp_conf = eotb_conf[1:15,]

# Don't show row names
row.names(disp_conf) = NULL

# Create table with gt package
conf = gt(disp_conf)
conf = conf %>%
  
  # Create title & Subtitle
  tab_header(
    title = 'Association Rules',
    subtitle = "Sorted by Confidence"
  ) %>%
  
  # Highlight Column of interest
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = Confidence)) %>%
  
  # Show High support values
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = Support,
      rows = Support > .1
    )
  )%>%
  
  # Show high lift values
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = Lift,
      rows = Lift > 2
    )
  )%>%
  
  # Bold source where both lift and support are high
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = Source,
      rows = (Support > .1 & Lift > 2)
  )) %>%
  
  # Bold target where both lift and support are high
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = Target,
      rows = (Support > .1 & Lift > 2)))

conf
gtsave(conf,'confidence_table.html')
rm(eotb_conf,conf)
```

## Creating Table of Support Values
```{r}

# Create dataframe with rules sorted by confidence
eotb_sup = DATAFRAME(eotb_rules_sup, separate = TRUE)

# Make rules more visually appealing
eotb_sup[] = lapply(eotb_sup, gsub, pattern='[{]', replacement='')
eotb_sup[] = lapply(eotb_sup, gsub, pattern='[}]', replacement='')
eotb_sup[] = lapply(eotb_sup, gsub, pattern='[_]', replacement=' ')
eotb_sup[] = lapply(eotb_sup, gsub, pattern='[,]', replacement=' & ')

# Get columns for table & change column names
eotb_sup = eotb_sup[,c('LHS','RHS','confidence','support','lift')]
colnames(eotb_sup) = c('Source','Target', 'Confidence','Support','Lift')

# Round numbers for table
eotb_sup[,3:5] = as.numeric(unlist(eotb_sup[,3:5]))
eotb_sup[,3:5] = round(eotb_sup[,3:5],3)

# Get top 15 rules
disp_sup = eotb_sup[1:15,]

# Don't show row names
row.names(disp_sup) = NULL

# Create table with gt package
sup = gt(disp_sup)
sup = sup %>%
  
  # Create title & Subtitle
  tab_header(
    title = 'Association Rules',
    subtitle = "Sorted by Support"
  ) %>%
  
  # Highlight Column of interest
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = Support)) %>%
  
  # Show High confidence values
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = Confidence,
      rows = Confidence > .75
    )
  )%>%
  
  # Show high lift values
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = Lift,
      rows = Lift > 1.8
    )
  )%>%
  
  # Bold source where both lift and confidence are high
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = Source,
      rows = (Confidence > .75 & Lift > 1.8)
  )) %>%
  
  # Bold target where both lift and confidence are high
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = Target,
      rows = (Confidence > .75 & Lift > 1.8)))

sup
gtsave(sup,'support_table.html')

rm(eotb_sup,sup)
```


## Creating Table of Lift Values
```{r}
# Create dataframe with rules sorted by confidence
eotb_lift = DATAFRAME(eotb_rules_lift, separate = TRUE)

# Make rules more visually appealing
eotb_lift[] = lapply(eotb_lift, gsub, pattern='[{]', replacement='')
eotb_lift[] = lapply(eotb_lift, gsub, pattern='[}]', replacement='')
eotb_lift[] = lapply(eotb_lift, gsub, pattern='[_]', replacement=' ')
eotb_lift[] = lapply(eotb_lift, gsub, pattern='[,]', replacement=' & ')

# Get columns for table & change column names
eotb_lift = eotb_lift[,c('LHS','RHS','confidence','support','lift')]
colnames(eotb_lift) = c('Source','Target', 'Confidence','Support','Lift')

# Round numbers for table
eotb_lift[,3:5] = as.numeric(unlist(eotb_lift[,3:5]))
eotb_lift[,3:5] = round(eotb_lift[,3:5],3)

# Get top 15 rules
disp_lift = eotb_lift[1:15,]

# Don't show row names
row.names(disp_lift) = NULL


# Create table with gt package
lift = gt(disp_lift)
lift = lift %>%
  
  # Create title & Subtitle
  tab_header(
    title = 'Association Rules',
    subtitle = "Sorted by Lift"
  ) %>%
  
  # Highlight Column of interest
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = Lift)) %>%
  
  # Show High confidence values
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = Confidence,
      rows = Confidence > .8
    )
  )%>%
  
  # Show high support values
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = Support,
      rows = Support > .1
    )
  )%>%
  
  # Bold source where both support and confidence are high
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = Source,
      rows = (Confidence > .8 & Support > .1)
  )) %>%
  
  # Bold target where both support and confidence are high
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = Target,
      rows = (Confidence > .8 & Support > .1)))

lift
gtsave(lift,'lift_table.html')
rm(eotb_lift,lift)
```

# Network D3 for Support
```{r}
# Create Edge List
edgeList = disp_sup[,c(1:2,4)]
(MyGraph = igraph::simplify(igraph::graph.data.frame(edgeList, directed=TRUE)))

# Create Node List
nodeList = data.frame(ID = c(0:(igraph::vcount(MyGraph) - 1)), 
                       nName = igraph::V(MyGraph)$name)

(nodeList = cbind(nodeList, nodeDegree=igraph::degree(MyGraph, 
                    v = igraph::V(MyGraph), mode = "all")))

## Get betweenness
BetweenNess = igraph::betweenness(MyGraph, 
      v = igraph::V(MyGraph), 
      directed = TRUE) 

(nodeList = cbind(nodeList, nodeBetweenness=BetweenNess))

# Build edges
getNodeID = function(x){
  which(x == igraph::V(MyGraph)$name) - 1  
}

edgeList = plyr::ddply(
  disp_sup, .variables = c("Source", "Target" , "Support"), 
  function (x) data.frame(SourceID = getNodeID(x$Source), 
                          TargetID = getNodeID(x$Target)))

# Find Dice Sim
DiceSim = igraph::similarity.dice(MyGraph, vids = igraph::V(MyGraph), mode = "all")


# New data frame with Dice similarity between vertices
F1 = function(x) {
  data.frame(diceSim = DiceSim[x$SourceID +1, x$TargetID + 1])
  }

# Put Dice Sim column in edgelist
head(edgeList)
edgeList = plyr::ddply(edgeList,
                        .variables=c("Source", "Target", "Support", 
                                               "SourceID", "TargetID"), 
                        function(x) data.frame(F1(x)))

# Create network 3D display
D3_network_eotb_support = networkD3::forceNetwork(
  Links = edgeList, # df info about edges
  Nodes = nodeList, # df info about nodes
  Source = "SourceID", # ID of source node 
  Target = "TargetID", # ID of target node
  Value = "Support", # from the edge list - value/weight relationship amongst nodes
  NodeID = "nName", # from the node list - node description 
  Nodesize = "nodeBetweenness",  # from the node list - node size
  Group = "nodeDegree",  # from the node list - node color
  height = 800, 
  width = 800,  
  fontSize = 20, 
  linkDistance = networkD3::JS("function(d) { return d.value*1000; }"), # Function to determine distance between any two nodes, uses variables already defined in forceNetwork function (not variables from a data frame)
  linkWidth = networkD3::JS("function(d) { return d.value*5; }"),# Function to determine link/edge thickness, uses variables already defined in forceNetwork function (not variables from a data frame)
  opacity = 5, 
  zoom = TRUE, 
  opacityNoHover = 7,
  linkColour = "red"   
) 

# Plot network
D3_network_eotb_support

# Save Network
networkD3::saveNetwork(D3_network_eotb_support, 
                       "d3_network_eotb_support.html", selfcontained = TRUE)
rm(D3_network_eotb_support,F1,edgelist,DiceSim,getNodeID,nodeList,BetweenNess,MyGraph)
```





# Network D3 for Confidence
```{r}
# Create Edge List
edgeList = disp_conf[,1:3]
(MyGraph = igraph::simplify(igraph::graph.data.frame(edgeList, directed=TRUE)))

nodeList = data.frame(ID = c(0:(igraph::vcount(MyGraph) - 1)), 
                       nName = igraph::V(MyGraph)$name)
## Create Node List
(nodeList = cbind(nodeList, nodeDegree=igraph::degree(MyGraph, 
                    v = igraph::V(MyGraph), mode = "all")))

## Get betweenness
BetweenNess = igraph::betweenness(MyGraph, 
      v = igraph::V(MyGraph), 
      directed = TRUE) 

(nodeList = cbind(nodeList, nodeBetweenness=BetweenNess))

# Build edges
getNodeID = function(x){
  which(x == igraph::V(MyGraph)$name) - 1  
}

edgeList = plyr::ddply(
  disp_conf, .variables = c("Source", "Target" , "Confidence"), 
  function (x) data.frame(SourceID = getNodeID(x$Source), 
                          TargetID = getNodeID(x$Target)))

# Find Dice Sim
DiceSim = igraph::similarity.dice(MyGraph, vids = igraph::V(MyGraph), mode = "all")


# New data frame with Dice similarity between vertices
F1 = function(x) {
  data.frame(diceSim = DiceSim[x$SourceID +1, x$TargetID + 1])
  }

# Put Dice Sim column in edgelist
head(edgeList)
edgeList = plyr::ddply(edgeList,
                        .variables=c("Source", "Target", "Confidence", 
                                               "SourceID", "TargetID"), 
                        function(x) data.frame(F1(x)))

# Create network 3D display
D3_network_eotb_confidence = networkD3::forceNetwork(
  Links = edgeList, # df info about edges
  Nodes = nodeList, # df info about nodes
  Source = "SourceID", # ID of source node 
  Target = "TargetID", # ID of target node
  Value = "Confidence", # from the edge list - value/weight relationship amongst nodes
  NodeID = "nName", # from the node list - node description 
  Nodesize = "nodeBetweenness",  # from the node list - node size
  Group = "nodeDegree",  # from the node list - node color
  height = 800, 
  width = 800,  
  fontSize = 20, 
  linkDistance = networkD3::JS("function(d) { return d.value*1000; }"), # Function to determine distance between any two nodes, uses variables already defined in forceNetwork function (not variables from a data frame)
  linkWidth = networkD3::JS("function(d) { return d.value*5; }"),# Function to determine link/edge thickness, uses variables already defined in forceNetwork function (not variables from a data frame)
  opacity = 5, 
  zoom = TRUE, 
  opacityNoHover = 7,
  linkColour = "red"   
) 

# Plot network
D3_network_eotb_confidence

# Save Network
networkD3::saveNetwork(D3_network_eotb_confidence, 
                       "d3_network_eotb_confidence.html", selfcontained = TRUE)
rm(D3_network_eotb_confidence,F1,edgelist,DiceSim,getNodeID,nodeList,BetweenNess,MyGraph)

```


# Network D3 for Lift
```{r}
# Create Edge List
edgeList = disp_lift[,c(1:2,5)]
(MyGraph = igraph::simplify(igraph::graph.data.frame(edgeList, directed=TRUE)))

nodeList = data.frame(ID = c(0:(igraph::vcount(MyGraph) - 1)), 
                       nName = igraph::V(MyGraph)$name)
## Create Node List
(nodeList = cbind(nodeList, nodeDegree=igraph::degree(MyGraph, 
                    v = igraph::V(MyGraph), mode = "all")))

## Get betweenness
BetweenNess = igraph::betweenness(MyGraph, 
      v = igraph::V(MyGraph), 
      directed = TRUE) 

(nodeList = cbind(nodeList, nodeBetweenness=BetweenNess))

# Build edges
getNodeID = function(x){
  which(x == igraph::V(MyGraph)$name) - 1  
}

edgeList = plyr::ddply(
  disp_lift, .variables = c("Source", "Target" , "Lift"), 
  function (x) data.frame(SourceID = getNodeID(x$Source), 
                          TargetID = getNodeID(x$Target)))

# Find Dice Sim
DiceSim = igraph::similarity.dice(MyGraph, vids = igraph::V(MyGraph), mode = "all")


# New data frame with Dice similarity between vertices
F1 = function(x) {
  data.frame(diceSim = DiceSim[x$SourceID +1, x$TargetID + 1])
  }

# Put Dice Sim column in edgelist
head(edgeList)
edgeList = plyr::ddply(edgeList,
                        .variables=c("Source", "Target", "Lift", 
                                               "SourceID", "TargetID"), 
                        function(x) data.frame(F1(x)))

# Create network 3D display
D3_network_eotb_lift = networkD3::forceNetwork(
  Links = edgeList, # df info about edges
  Nodes = nodeList, # df info about nodes
  Source = "SourceID", # ID of source node 
  Target = "TargetID", # ID of target node
  Value = "Lift", # from the edge list - value/weight relationship amongst nodes
  NodeID = "nName", # from the node list - node description 
  Nodesize = "nodeBetweenness",  # from the node list - node size
  Group = "nodeDegree",  # from the node list - node color
  height = 800, 
  width = 800,  
  fontSize = 20, 
  linkDistance = networkD3::JS("function(d) { return d.value*1000; }"), # Function to determine distance between any two nodes, uses variables already defined in forceNetwork function (not variables from a data frame)
  linkWidth = networkD3::JS("function(d) { return d.value*5; }"),# Function to determine link/edge thickness, uses variables already defined in forceNetwork function (not variables from a data frame)
  opacity = 5, 
  zoom = TRUE, 
  opacityNoHover = 7,
  linkColour = "red"   
) 

# Plot network
D3_network_eotb_lift

# Save Network
networkD3::saveNetwork(D3_network_eotb_lift, 
                       "d3_network_eotb_lift.html", selfcontained = TRUE)
rm(D3_network_eotb_lift,F1,edgelist,DiceSim,getNodeID,nodeList,BetweenNess,MyGraph)

```

## 3D plot of rule measurements
```{r}
# Create Data frame of rules
eotb_rules_df = DATAFRAME(eotb_rules, separate = TRUE)

# x-axis title
axx = list(
  title = "Support"
)

# y-axis title
axy = list(
  title = "Confidence"
)

# z-axis title
axz = list(
  title = "Lift"
)

# Change colors for high values of each measurement
eotb_rules_df$color = ' '
eotb_rules_df$color[eotb_rules_df$support > .2] = 'High Support'
eotb_rules_df$color[eotb_rules_df$confidence > .9] = 'High Confidence'
eotb_rules_df$color[eotb_rules_df$lift > 4.5] = 'High Lift'
eotb_rules_df$color[eotb_rules_df$lift > 4.5 & eotb_rules_df$confidence > .9] = 'High Confidence & Lift'


# Create 3D visual with plotly
fig = plot_ly(eotb_rules_df, x = ~support, y = ~confidence, z = ~lift)
fig = fig %>% add_markers(color = ~color)
fig = fig %>% layout(title = '3D Plot of Association Rules',
                      scene = list(
                        xaxis = axx,
                        yaxis = axy,
                        zaxis = axz
                      ))
# View figure
fig


# Save as html widget
saveWidget(fig, "Association_Rules_3d_plot.html")
rm(fig,eotb_rules_df,axx,axy,axz)

```

## Parallel Coordinates Plot - Support
```{r}
png('par_coord_sup.png',width=1000, height=600)
plot(eotb_rules_sup[1:15], method="paracoord", control=list(alpha=0.8, reorder=TRUE), main = 'Parallel Coordinates Plot for Top 15 Support Items')
dev.off()

```

## Parallel Coordinates Plot - Confidence
```{r}
png('par_coord_conf.png',width=1000, height=600)

plot(eotb_rules_conf[1:15], 
     method="paracoord", 
     control=list(alpha=.9, reorder=TRUE), 
     main = 'Parallel Coordinates Plot for Top 15 Confidence Items')

dev.off()

```

## Parallel Coordinates Plot - Lift
```{r}
png('par_coord_lift.png',width=1000, height=600)

plot(eotb_rules_lift[1:15], 
     method="paracoord", 
     control=list(alpha=0.9, reorder=TRUE), 
     main = 'Parallel Coordinates Plot for Top 15 Lift Items')

dev.off()

```



