---
title: "R Notebook"
output: html_notebook
---

## Import Libraries
```{r}

library(tidyverse)
library(ggplot2)
library(rpart)
library(rattle) 
library(rpart.plot)
library(RColorBrewer)
library(network)
library(plyr)
library(factoextra)
library(GGally)
library(ggpubr)
library(lubridate)
library(plotly) 
library(htmlwidgets)
library(philentropy)
library(gt)
library(igraph)
library(cvms)
library(tibble)
library(caret)
library(gridExtra)     
library(caretEnsemble)
library(tm)
library(slam)
library(quanteda)
library(SnowballC)
library(arules)
library(proxy)
library(cluster)
library(Matrix)
library(plyr) 
library(mclust) 
library(naivebayes)
library(psych)
library(mice)
library(e1071)
library(mlr)
```


## Confusion Matrix Function
```{r}
### Function below from: https://stackoverflow.com/a/66606295 for confusion matrix visual

conf_matrix <- function(df.true, df.pred, true.lab ="True Class", pred.lab ="Predicted Class",
                        high.col = 'green3', low.col = 'white',title = "Confusion Matrix",subtit = "") {
  #convert input vector to factors, and ensure they have the same levels
  df.true <- as.factor(df.true)
  df.pred <- factor(df.pred, levels = levels(df.true))
  
  #generate confusion matrix, and confusion matrix as a pecentage of each true class (to be used for color) 
  df.cm <- table(True = df.true, Pred = df.pred)
  df.cm.col <- df.cm / rowSums(df.cm)
  
  #convert confusion matrices to tables, and binding them together
  df.table <- reshape2::melt(df.cm)
  df.table.col <- reshape2::melt(df.cm.col)
  df.table <- left_join(df.table, df.table.col, by =c("True", "Pred"))
  
  #calculate accuracy and class accuracy
  acc.vector <- c(diag(df.cm)) / c(rowSums(df.cm))
  class.acc <- data.frame(Pred = "Class Acc.", True = names(acc.vector), value = acc.vector)
  acc <- sum(diag(df.cm)) / sum(df.cm)
  
  #plot
  ggplot() +
    geom_tile(aes(x=Pred, y=True, fill=value.y),
              data=df.table, size=0.2, color=grey(0.5)) +
    geom_tile(aes(x=Pred, y=True),
              data=df.table[df.table$True==df.table$Pred, ], size=1, color="black", fill = 'transparent') +
    scale_x_discrete(position = "top",  limits = c(levels(df.table$Pred), "Class Acc.")) +
    scale_y_discrete(limits = rev(unique(levels(df.table$Pred)))) +
    labs(x=pred.lab, 
         y=true.lab, 
         fill=NULL,
         title = title,
         subtitle = subtit,
         caption = paste0("Total Accuracy: ", round(100*acc, 1), "%")) +
    geom_text(aes(x=Pred, y=True, label=value.x),
              data=df.table, size=4, colour="black") +
    geom_text(data = class.acc, aes(Pred, True, label = paste0(round(100*value), "%"))) +
    scale_fill_gradient(low=low.col, high=high.col, labels = scales::percent,
                        limits = c(0,1), breaks = c(0,0.5,1)) +
    guides(scale = "none") +
    theme_bw() +
    theme(panel.border = element_blank(), legend.position = "bottom",
          axis.text = element_text(color='black'), axis.ticks = element_blank(),
          panel.grid = element_blank(), axis.text.x.top = element_text(angle = 30, vjust = 0, hjust = 0),
          plot.title = element_text(size=17, face="bold", 
      margin = margin(10, 0, 10, 0),hjust =.5),
      plot.subtitle = element_text(size = 14, face = 'bold',hjust =.5)) +
    coord_fixed()

} 

```

## Read in Data
```{r}
eotb = read.csv('EOTB_Cleaned.csv')
eotb$DateTime = lubridate::parse_date_time(eotb$DateTime, "ymd HMS")

# View data structure again
#str(eotb)
# Structure looks good now
```

## Subsetting data for SVM
```{r}
# Need only numeric data so will subset the data. 
# Due to size of data, subsetting dataset for just the span of a year 
num_eotb = eotb[(year(eotb$DateTime) > 2010 )& (year(eotb$DateTime) < 2012 ),]

```

## Clean & Structure Data
```{r}
# Get Rid of non numeric/factor columns
num_eotb = num_eotb[,5:10]

# Rename column
colnames(num_eotb) = c("Salinity","pH","Dis_Oxygen","Turbidity","Chlorophyll","Temperature")
#eotb$Station = as.factor(eotb$Station)
#eotb$Month = as.factor(eotb$Month)
```



## Change Oxygen Values to labels
```{r}
# Change Oxygen Values to labels - target of study for now. 
ox_lab = num_eotb

do = ox_lab[,3]

tdo = function(do){
  if(do < 70)
    x = 'Low Oxygen'
  else
    x = "Not Low Oxygen"
  return(x)
}
dissolved_oxygen = sapply(do, function(x) tdo(x))
ox_lab$Dis_Oxygen = as.factor(dissolved_oxygen)
rm(do,tdo,dissolved_oxygen)

```


## View amounts of each label
```{r}
#table(ox_lab$Dis_Oxygen)

```

## Scale the data
```{r}
ox_lab$Salinity = scale(ox_lab$Salinity)
ox_lab$pH = scale(ox_lab$pH)
ox_lab$Turbidity = scale(ox_lab$Turbidity)
ox_lab$Chlorophyll = scale(ox_lab$Chlorophyll)
ox_lab$Temperature = scale(ox_lab$Temperature)

```

## Split into test & training set
```{r}
set.seed = 123
# Need even amounts in training set to train data:, low_oxygen has fewest labels so least common denominator

# Create column of index rows
ox_lab$row_num = 1:nrow(ox_lab)

# Find amount of low oxygen values
l = nrow(ox_lab[ox_lab$Dis_Oxygen == 'Low Oxygen',])

# Make training size 70% of set
train_size = l*.7

# Extract indicies of even amount of each label for training
low_o_sam = sample(ox_lab$row_num[ox_lab$Dis_Oxygen == 'Low Oxygen'],train_size)
not_low_o_sam = sample(ox_lab$row_num[ox_lab$Dis_Oxygen == 'Not Low Oxygen'],train_size)

# Combine Indicies
train_ind = c(not_low_o_sam,low_o_sam)

# Remove row indicies from df
ox_lab = ox_lab[,1:6]

# Get training and Testing Set
train_set = ox_lab[train_ind,]
test_set = ox_lab[-train_ind,]

# Save Labels in variables
test_labs = test_set$Dis_Oxygen
train_labs = train_set$Dis_Oxygen

# Remove Labels from sets
test_set = test_set[,c(1:2,4:6)]
train_set = train_set[,c(1:2,4:6)]
#rm(not_low_o_sam,low_o_sam,train_ind)
```

# Run SVM on many different kernels
```{r}
# Set vectors for tracking accuracy
acc = c()
svm_type = c('linear','quadratic','cubic','quartic','radial','sigmoid')
```



```{r}
# Linear SVM

l_svm = e1071::svm(train_labs~., data=train_set, 
            kernel="linear",
            scale = F)

pred = predict(l_svm, test_set, type="class")

tb = table(pred,test_labs)
acc = append(acc,sum(diag(tb))/sum(tb))


# Quadratic SVM
qd_svm = e1071::svm(train_labs~., data=train_set, 
            kernel="polynomial", 
            degree = 2)

pred = predict(qd_svm, test_set, type="class")

tb = table(pred,test_labs)
acc = append(acc,sum(diag(tb))/sum(tb))

# Cubic SVM
c_svm = e1071::svm(train_labs~., data=train_set, 
            kernel="polynomial", 
            degree = 3)

pred = predict(c_svm, test_set, type="class")

tb = table(pred,test_labs)
acc = append(acc,sum(diag(tb))/sum(tb))

# Quartic SVM
qr_svm = e1071::svm(train_labs~., data=train_set, 
            kernel="polynomial", 
            degree = 4
)

pred = predict(qr_svm, test_set, type="class")

tb = table(pred,test_labs)
acc = append(acc,sum(diag(tb))/sum(tb))

# Radial SVM
rad_svm = e1071::svm(train_labs~., data=train_set, 
            kernel="radial")

pred = predict(rad_svm, test_set, type="class")

tb = table(pred,test_labs)
acc = append(acc,sum(diag(tb))/sum(tb))

# Sigmoid SVM
sig_svm = e1071::svm(train_labs~., data=train_set, 
            kernel="sigmoid")

pred = predict(sig_svm, test_set, type="class")

tb = table(pred,test_labs)
acc = append(acc,sum(diag(tb))/sum(tb))


acc_df = data.frame(svm_type,round(acc,4))

colnames(acc_df) = c('SVM Type','Accuracy')
acc_df
## Radial has highest accuracy.
```

# Save accuracy table
```{r}
# Create table
acc_tb = gt(acc_df)
acc_tb = acc_tb %>%
  
# Create title & Subtitle
  tab_header(
    title = 'Accuracy of Different SVM Kernels for Dissolved Oxygen Labels')%>%
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = Accuracy,
      rows = Accuracy > 0.8
    ))
acc_tb
gtsave(acc_tb,'OX_acc_tab.html')


```

## Try Radial Kernal with different costs 
```{r}
acc = c()
costs = c(0.1,0.5,1,2.5,5,10,25,50,100)
for (c in costs){
  start = Sys.time()
  rad_svm = e1071::svm(train_labs~., 
                       data=train_set,
                       cost = c,
                       kernel="radial")

  pred = predict(rad_svm, test_set, type="class")
  end = Sys.time()
  start - end
  tb = table(pred,test_labs)
  acc = append(acc,sum(diag(tb))/sum(tb))
}
```
```{r}

acc_df = data.frame(costs,acc)
colnames(acc_df) = c('Cost','Accuracy')
acc_df
ggplot(data = acc_df)+
  geom_line(aes(Cost,Accuracy),color = 'blue')+
  geom_point(aes(Cost,Accuracy),color = 'red')+
  geom_text(aes(label = Cost,x =Cost, y = Accuracy),color = 'black', position=position_dodge(width=0.9),hjust = 1.5, vjust=-.5,size =3)+
  labs(title = 'SVM Accuracy Gain in Relation to Cost')+
  theme(plot.title = element_text(size=17, face="bold", 
      margin = margin(10, 0, 10, 0),hjust =.5) )
ggsave('Accuracy_gain_ox.png')
 
```
```{r}
start = Sys.time()
rad_svm = e1071::svm(train_labs~., 
                     data=train_set,
                     cost = 2.5,
                     kernel="radial")

pred = predict(rad_svm, test_set, type="class")
end = Sys.time()
start - end



```

# Confusion Matrix & SVM Classification Plot
```{r}
# COnfusion Matrix
cm = conf_matrix(test_labs,pred,title = 'Confusion Matrix for SVM model of Dissolved Oxygen')
cm
ggsave('ox_cm.png',plot = cm)

# Classification Plot
png(file="svm_class_plot_OX.png")
plot(rad_svm, data=train_set, pH~Temperature,
       slice = list(Salinity = 0,Turbidity=0,Chlorophyll=0),
     xlim = c(-2.5,2.25),
     ylim = c(-3.9,2.9),
     symbolPalette = rainbow(3),
)
dev.off()


```

## Test on more data
```{r}
# Get data from 2016-2020
pres_eotb = eotb[(year(eotb$DateTime) > 2016 )& (year(eotb$DateTime) <= 2020 ),]

# Clean same way as other data
pres_eotb = pres_eotb[,5:10]
colnames(pres_eotb) = c("Salinity","pH","Dis_Oxygen","Turbidity","Chlorophyll","Temperature")

ox_lab = num_eotb

do = ox_lab[,3]

tdo = function(do){
  if(do < 70)
    x = 'Low Oxygen'
  else
    x = "Not Low Oxygen"
  return(x)
}
dissolved_oxygen = sapply(do, function(x) tdo(x))
ox_lab$Dis_Oxygen = as.factor(dissolved_oxygen)
rm(do,tdo,dissolved_oxygen)

# Normalize
ox_lab$Salinity = scale(ox_lab$Salinity)
ox_lab$pH = scale(ox_lab$pH)
ox_lab$Turbidity = scale(ox_lab$Turbidity)
ox_lab$Chlorophyll = scale(ox_lab$Chlorophyll)
ox_lab$Temperature = scale(ox_lab$Temperature)

# save and remove labels
pres_lab = ox_lab$Dis_Oxygen
ox_lab = ox_lab[,c(1:2,4:6)]

# Run Prediction
pred = predict(rad_svm, ox_lab, type="class")

# Confusion Matrix
cm = conf_matrix(pres_lab,pred,title = 'Confusion Matrix for SVM model of Dissolved Oxygen')
cm
ggsave('ox_cm_pres.png',plot = cm)



```


# Next do chlorophyll levels


## Change chlorophyll Values to labels
```{r}

cl_lab = num_eotb

cl = cl_lab[5]
#summary(cl)
cl = cl[cl > 0 & cl < 500]

tcl = function(cl){
  if(cl < 15)
    x  = 'Not Harmful CL'

  else if(cl >= 15)
    x = 'Harmful CL'
  return(x)
}
cl = sapply(cl, function(x) tcl(x))
cl_lab$Chlorophyll = as.factor(cl)
rm(cl,tcl)

```


## Scale the data
```{r}
cl_lab$Salinity = scale(cl_lab$Salinity)
cl_lab$pH = scale(cl_lab$pH)
cl_lab$Turbidity = scale(cl_lab$Turbidity)
cl_lab$Dis_Oxygen = scale(cl_lab$Dis_Oxygen)
cl_lab$Temperature = scale(cl_lab$Temperature)

```

## Split into test & training set
```{r}
set.seed = 123
# Need even amounts in training set to train data:, low_oxygen has fewest labels so least common denominator

# Create column of index rows
cl_lab$row_num = 1:nrow(cl_lab)

# Find amount of low oxygen values
l = nrow(cl_lab[cl_lab$Chlorophyll == 'Harmful CL',])

# Make training size 70% of set
train_size = l*.7

# Extract indicies of even amount of each label for training
harm_sam = sample(cl_lab$row_num[cl_lab$Chlorophyll == 'Harmful CL'],train_size)
not_harm_sam = sample(cl_lab$row_num[cl_lab$Chlorophyll == 'Not Harmful CL'],train_size)

# Combine Indicies
train_ind = c(not_harm_sam,harm_sam)

# Remove row indicies from df
cl_lab = cl_lab[,1:6]

# Get training and Testing Set
train_set = cl_lab[train_ind,]
test_set = cl_lab[-train_ind,]

# Save Labels in variables
test_labs = test_set$Chlorophyll
train_labs = train_set$Chlorophyll

# Remove Labels from sets
test_set = test_set[,c(1:4,6)]
train_set = train_set[,c(1:4,6)]
#rm(not_low_o_sam,low_o_sam,train_ind)
```

# Run SVM on many different kernels
```{r}
# Set vectors for tracking accuracy
acc = c()
svm_type = c('linear','quadratic','cubic','quartic','radial','sigmoid')
```



```{r}
# Linear SVM
Start = Sys.time()
l_svm = e1071::svm(train_labs~., data=train_set, 
            kernel="linear",
            scale = F)

pred = predict(l_svm, test_set, type="class")

tb = table(pred,test_labs)
acc = append(acc,sum(diag(tb))/sum(tb))


# Quadratic SVM
qd_svm = e1071::svm(train_labs~., data=train_set, 
            kernel="polynomial", 
            degree = 2)

pred = predict(qd_svm, test_set, type="class")

tb = table(pred,test_labs)
acc = append(acc,sum(diag(tb))/sum(tb))

# Cubic SVM
c_svm = e1071::svm(train_labs~., data=train_set, 
            kernel="polynomial", 
            degree = 3)

pred = predict(c_svm, test_set, type="class")

tb = table(pred,test_labs)
acc = append(acc,sum(diag(tb))/sum(tb))

# Quartic SVM
qr_svm = e1071::svm(train_labs~., data=train_set, 
            kernel="polynomial", 
            degree = 4
)

pred = predict(qr_svm, test_set, type="class")

tb = table(pred,test_labs)
acc = append(acc,sum(diag(tb))/sum(tb))

# Radial SVM
rad_svm = e1071::svm(train_labs~., data=train_set, 
            kernel="radial")

pred = predict(rad_svm, test_set, type="class")

tb = table(pred,test_labs)
acc = append(acc,sum(diag(tb))/sum(tb))

# Sigmoid SVM
sig_svm = e1071::svm(train_labs~., data=train_set, 
            kernel="sigmoid")

pred = predict(sig_svm, test_set, type="class")

tb = table(pred,test_labs)
acc = append(acc,sum(diag(tb))/sum(tb))

end = Sys.time()
end-start

acc_df = data.frame(svm_type,round(acc,4))

colnames(acc_df) = c('SVM Type','Accuracy')
acc_df
## quartic has highest accuracy.
```
# Save accuracy table
```{r}
# Create table
acc_tb = gt(acc_df)
acc_tb = acc_tb %>%
  
# Create title & Subtitle
  tab_header(
    title = 'Accuracy of Different SVM Kernels for Chlorophyll Labels')%>%
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = Accuracy,
      rows = Accuracy > 0.88
    ))
acc_tb
gtsave(acc_tb,'CL_acc_tab.html')


```
## Try Radial Kernal with different costs 
```{r}
rad_acc = c()
costs = c(0.1,0.5,1,2.5,5,10,25,50,100)
for (c in costs){
  rad_svm = e1071::svm(train_labs~., 
                       data=train_set,
                       cost = c,
                       kernel="radial")

  rad_pred = predict(rad_svm, test_set, type="class")
  rad_tb = table(rad_pred,test_labs)
  rad_acc = append(rad_acc,sum(diag(rad_tb))/sum(rad_tb))
}
```

## Try Quartic Kernal with different costs 
```{r}
qrt_acc = c()
costs = c(0.1,0.5,1,2.5,5,10,25,50,100)
for (c in costs){
  qrt_svm = e1071::svm(train_labs~., 
                       data=train_set,
                       cost = c,
                       kernel="polynomial",
                       degree=4)

  qrt_pred = predict(qrt_svm, test_set, type="class")
  qrt_tb = table(qrt_pred,test_labs)
  qrt_acc = append(qrt_acc,sum(diag(qrt_tb))/sum(qrt_tb))
}
```

## Plot cost vs accuracy
```{r}

rad_acc_df = data.frame(costs,rad_acc)
colnames(rad_acc_df) = c('Cost','Accuracy')
acc_df
ggplot(data = rad_acc_df)+
  geom_line(aes(Cost,Accuracy),color = 'blue')+
  geom_point(aes(Cost,Accuracy),color = 'red')+
  geom_text(aes(label = Cost,x =Cost, y = Accuracy),color = 'black', position=position_dodge(width=0.9), vjust=-.5,size =3)+
  labs(title = 'SVM Accuracy Gain in Relation to Cost',
       subtitle = 'Radial Kernel')+
  theme(plot.title = element_text(size=17, face="bold", 
      margin = margin(10, 0, 10, 0),hjust =.5) )
ggsave('Accuracy_gain_cl_radial.png')
```
```{r}

qrt_acc_df = data.frame(costs,qrt_acc)
colnames(qrt_acc_df) = c('Cost','Accuracy')
qrt_acc_df
ggplot(data = qrt_acc_df)+
  geom_line(aes(Cost,Accuracy),color = 'blue')+
  geom_point(aes(Cost,Accuracy),color = 'red')+
  geom_text(aes(label = Cost,x =Cost, y = Accuracy),color = 'black', position=position_dodge(width=0.9), vjust=-.5,size =3)+
  labs(title = 'SVM Accuracy Gain in Relation to Cost',
       subtitle = 'Quartic Kernel')+
  theme(plot.title = element_text(size=17, face="bold", 
      margin = margin(10, 0, 10, 0),hjust =.5) )
ggsave('Accuracy_gain_cl_quartic.png')
```

## Run on radial kernel with 2.5 cost
```{r}

rad_svm = e1071::svm(train_labs~., 
                     data=train_set,
                     cost = 2.5,
                     kernel="radial")

rad_pred = predict(rad_svm, test_set, type="class")




```
## Run on quartic kernel with 0.1 cost
```{r}

qrt_svm = e1071::svm(train_labs~., 
                     data=train_set,
                     cost = 0.1,
                     kernel="polynomial",
                     degree = 4)

qrt_pred = predict(qrt_svm, test_set, type="class")




```


## Plot confusion matrix & SVM classification plot
```{r}
# Confusion Matrix
cm = conf_matrix(test_labs,rad_pred,title = 'Confusion Matrix for SVM model of Chlorophyll')
cm
ggsave('cl_cm_rad.png',plot = cm)

# Classification Plot
png(file="svm_class_plot_CL_rad.png")
plot(rad_svm, data=train_set, Salinity~Temperature,
       slice = list(pH = 0,Turbidity=0,Dis_Oxygen=0),
     xlim = c(-2.25,2.1),
     ylim = c(-.8,2.25),
     symbolPalette = rainbow(3),
)

dev.off()


```
```{r}
# Confusion Matrix
cm = conf_matrix(test_labs,qrt_pred,title = 'Confusion Matrix for SVM model of Chlorophyll')
cm
ggsave('cl_cm_qrt.png',plot = cm)

# Classification Plot
png(file="svm_class_plot_CL_qrt.png")
plot(qrt_svm, data=train_set, Salinity~Temperature,
       slice = list(pH = 0,Turbidity=0,Dis_Oxygen=0),
     xlim = c(-2.5,2.1),
     ylim = c(-.9,2.25),
     symbolPalette = rainbow(3),
)

dev.off()


```

## Test on more data
```{r}
# Get data from 2016-2020
pres_eotb = eotb[(year(eotb$DateTime) > 2016 )& (year(eotb$DateTime) <= 2020 ),]

# Clean same way as other data
pres_eotb = pres_eotb[,5:10]
colnames(pres_eotb) = c("Salinity","pH","Dis_Oxygen","Turbidity","Chlorophyll","Temperature")

cl_lab = pres_eotb

cl = cl_lab[5]
#summary(cl)
cl = cl[cl > 0 & cl < 500]

tcl = function(cl){
  if(cl < 15)
    x  = 'Not Harmful CL'

  else if(cl >= 15)
    x = 'Harmful CL'
  return(x)
}
cl = sapply(cl, function(x) tcl(x))
cl_lab$Chlorophyll = as.factor(cl)
rm(cl,tcl)


# Normalize
cl_lab$Salinity = scale(cl_lab$Salinity)
cl_lab$pH = scale(cl_lab$pH)
cl_lab$Turbidity = scale(cl_lab$Turbidity)
cl_lab$Dis_Oxygen = scale(cl_lab$Dis_Oxygen)
cl_lab$Temperature = scale(cl_lab$Temperature)

# save and remove labels
pres_lab = cl_lab$Chlorophyll
cl_lab = cl_lab[,c(1:4,6)]

# Run Prediction
new_rad_pred = predict(rad_svm, cl_lab, type="class")
new_qrt_pred = predict(qrt_svm, cl_lab, type="class")



# Confusion Matrices
cm = conf_matrix(pres_lab,new_rad_pred,title = 'Confusion Matrix for SVM model of Chlorophyll',subtit = 'Radial Kernel')
cm
ggsave('cl_cm_pres_rad.png',plot = cm)


cm = conf_matrix(pres_lab,new_qrt_pred,title = 'Confusion Matrix for SVM model of Chlorophyll',subtit = 'Quartic Kernel')
cm
ggsave('cl_cm_pres_qrt.png',plot = cm)



```
