---
title: "R Notebook"
output: html_notebook
---

## Libraries
```{r}
library(ggplot2)
library(XLConnect)
library(tidyverse)
library(ggimage)
library(jsonlite)
library(httr)
library(rvest)
library(gghighlight)
library(reshape2) 
library(kableExtra)
library(ggpubr)
library(gt)
`%!in%` <- Negate(`%in%`)
options(java.parameters = "- Xmx1024m")
```

## Import cleaned salary dataset
```{r}
sal = read_csv('all_seasons_salaries.csv')

```
## Match Players to teams by season & Add statistics columns
```{r}
# Amount of rows in data set
seasons = 12
teams = 30
l = seasons*teams

# Copy dataset
copy_sal = sal

# Create empty new dataframe for output
sal_dist = data.frame('Team'=character(l),
                      'Season'=character(l),
                      'Mean_Salary'=numeric(l),
                      'Median-Salary'=numeric(l),
                      'Max'=numeric(l),
                      'Min'=numeric(l),
                      'Standard_deviation'=numeric(l),
                      'Total' = numeric(l),
                      'salaries' = numeric(l))

# For counting players on teams
i = 1

# For loop to transform data to new dataframe
for(z in 1:l){
  
  # Get team and season
  team = copy_sal$TEAM[1]
  yr = copy_sal$year[1]
  
  # Extract salaries by team and season
  sals = copy_sal$salary[copy_sal$TEAM == team & copy_sal$year == yr & !is.na(copy_sal$salary)]
  
  # Place team, season, & salary statistics in data frame
  sal_dist$Team[z] = team
  sal_dist$Season[z] = yr
  sal_dist$Mean_Salary[z] = mean(sals)
  sal_dist$Median.Salary[z] = median(sals)
  sal_dist$Max[z] = max(sals)
  sal_dist$Min[z] = min(sals)
  sal_dist$Standard_deviation[z] = sd(sals)
  sal_dist$Total[z] = sum(sals)
  sal_dist$salaries[z] = list(sals)
  
  # Remove data from copied data frame
  c_len = length(copy_sal$salary)
  len = length(sals)
  copy_sal = copy_sal[1+len:c_len,]

}
```


# Explore Data using box plots
```{r}
boxplot(sal_dist$Mean_Salary)
boxplot(sal_dist$Median.Salary)
boxplot(sal_dist$Max)
boxplot(sal_dist$Min)
boxplot(sal_dist$Standard_deviation)
boxplot(sal_dist$Total)




```
## Match abbreviation to city to match dataset with win dataset
```{r}

for(i in 1:length(sal_dist$Team)){
  if(sal_dist$Team[i] == "ATL")
    sal_dist$nickname[i] = "Atlanta"
  else if(sal_dist$Team[i] == "BOS")
    sal_dist$nickname[i] = "Boston"
  else if(sal_dist$Team[i] == "CHA")
    sal_dist$nickname[i] = "Charlotte"
  else if(sal_dist$Team[i] == "CHI")
    sal_dist$nickname[i] = "Chicago"
  else if(sal_dist$Team[i] == "CLE")
    sal_dist$nickname[i] = "Cleveland"
  else if(sal_dist$Team[i] == "DAL")
    sal_dist$nickname[i] = "Dallas"
  else if(sal_dist$Team[i] == "DEN")
    sal_dist$nickname[i] = "Denver"
  else if(sal_dist$Team[i] == "DET")
    sal_dist$nickname[i] = "Detroit"
  else if(sal_dist$Team[i] == "GSW")
    sal_dist$nickname[i] = "Golden State"
  else if(sal_dist$Team[i] == "HOU")
    sal_dist$nickname[i] = "Houston"
  else if(sal_dist$Team[i] == "IND")
    sal_dist$nickname[i] = "Indiana"
  else if(sal_dist$Team[i] == "LAC")
    sal_dist$nickname[i] = "LA Clippers"
  else if(sal_dist$Team[i] == "LAL")
    sal_dist$nickname[i] = "LA Lakers"
  else if(sal_dist$Team[i] == "MEM")
    sal_dist$nickname[i] = "Memphis"
  else if(sal_dist$Team[i] == "MIA")
    sal_dist$nickname[i] = "Miami"
  else if(sal_dist$Team[i] == "MIL")
    sal_dist$nickname[i] = "Milwaukee"
  else if(sal_dist$Team[i] == "MIN")
    sal_dist$nickname[i] = "Minnesota"
  else if(sal_dist$Team[i] == "NJN")
    sal_dist$nickname[i] = "New Jersey"
  else if(sal_dist$Team[i] == "NOH")
    sal_dist$nickname[i] = "New Orleans"
  else if(sal_dist$Team[i] == "NYK")
    sal_dist$nickname[i] = "New York"
  else if(sal_dist$Team[i] == "OKC")
    sal_dist$nickname[i] = "Okla City"
  else if(sal_dist$Team[i] == "ORL")
    sal_dist$nickname[i] = "Orlando"
  else if(sal_dist$Team[i] == "PHI")
    sal_dist$nickname[i] = "Philadelphia"
  else if(sal_dist$Team[i] == "PHX")
    sal_dist$nickname[i] = "Phoenix"
  else if(sal_dist$Team[i] == "POR")
    sal_dist$nickname[i] = "Portland"
  else if(sal_dist$Team[i] == "SAC")
    sal_dist$nickname[i] = "Sacramento"
  else if(sal_dist$Team[i] == "SAS")
    sal_dist$nickname[i] = "San Antonio"
  else if(sal_dist$Team[i] == "TOR")
    sal_dist$nickname[i] = "Toronto"
  else if(sal_dist$Team[i] == "UTA")
    sal_dist$nickname[i] = "Utah"
  else if(sal_dist$Team[i] == "WAS")
    sal_dist$nickname[i] = "Washington"
  else if(sal_dist$Team[i] == "BKN")
    sal_dist$nickname[i] = "Brooklyn"
  else if(sal_dist$Team[i] == "NOP")
    sal_dist$nickname[i] = "New Orleans"
  
}
```

## Clean up naming & column order of new dataset
```{r}
names(sal_dist)[10] = "City"
sal_dist = sal_dist[,c("Team","City","Season","Mean_Salary","Median.Salary",
                      "Max","Min","Standard_deviation","Total","salaries")]
```

## Combine dataset with Wins dataset
```{r}
wins = readWorksheetFromFile("wins.xlsx",sheet = 1,header = T)
sal_dist = sal_dist %>%
  left_join(wins, by = c('City' = 'Team','Season'='year'))
#sum(is.na(sal_dist$wp))
```

```{r}
# Save in case needed for later usage
write_csv(sal_dist,'sal_dist.csv')

```

## View Mean differences between winning and losing teams. 
```{r}
# Split winning and losing team labels
sal_dist$WinTeam = "Winning Teams"
sal_dist$WinTeam[sal_dist$wp < .5] = "Losing Teams"

# Save as variables
w = sal_dist$WinTeam[sal_dist$WinTeam == "Winning Teams"]
l = sal_dist$WinTeam[sal_dist$WinTeam == "Losing Teams"]

# NBA colors
logocolors = c('#17408B',"#C9082A")

# Boxplot
box_mean = ggplot(data = sal_dist)+
  geom_boxplot(aes(x = WinTeam,y = Mean_Salary),color = logocolors)+
  labs(x = "",
       y = "Mean Player Salary",
       title = "Mean Player Salaries by Winning & Losing Teams",
       subtitle = "2009-2020 Seasons"
         ) +
  theme(plot.title = element_text(size=17, face="bold", 
    margin = margin(10, 0, 10, 0),hjust =.5),
    plot.subtitle = element_text(hjust =.5),
    axis.text=element_text(size=10),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank())+
  scale_y_continuous(labels = scales::comma)
box_mean
ggsave('mean_box.png',width = 12, height = 9, dpi = 'retina')

  
```
## View Median differences between winning and losing teams. 
```{r}

box_median = ggplot(data = sal_dist)+
  geom_boxplot(aes(x = WinTeam,y = Median.Salary),color = logocolors)+
  labs(x = "",
       y = "Median Player Salary",
       title = "Median Player Salaries by Winning & Losing Teams",
       subtitle = "2009-2020 Seasons"
         ) +
  theme(plot.title = element_text(size=17, face="bold", 
    margin = margin(10, 0, 10, 0),hjust =.5),
    plot.subtitle = element_text(hjust =.5),
    axis.text=element_text(size=10),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank())+
  scale_y_continuous(labels = scales::comma)
box_median
ggsave('median_box.png',width = 12, height = 9, dpi = 'retina')

  
```
## View Standard deviation differences between winning and losing teams. 
```{r}

box_sd = ggplot(data = sal_dist)+
  geom_boxplot(aes(x = WinTeam,y = Standard_deviation),color = logocolors)+
  labs(x = "",
       y = "Standard Deviation",
       title = "Standard Deviation of Team's Player Salaries",
       subtitle = "By Winning & Losing Teams (2009-2020 Seasons)"
         ) +
  theme(plot.title = element_text(size=17, face="bold", 
    margin = margin(10, 0, 10, 0),hjust =.5),
    plot.subtitle = element_text(hjust =.5),
    axis.text=element_text(size=10),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank())+
  scale_y_continuous(labels = scales::comma)
box_sd
ggsave('sd_box.png',width = 12, height = 9, dpi = 'retina')

  
```
## View Total Spend differences between winning and losing teams. 
```{r}

box_total = ggplot(data = sal_dist)+
  geom_boxplot(aes(x = WinTeam,y = Total),color = logocolors)+
  labs(x = "",
       y = "Total Salary Spent",
       title = "Total Salaray Spent by Winning & Losing Teams",
       subtitle = "2009-2020 Seasons"
         ) +
  theme(plot.title = element_text(size=17, face="bold", 
    margin = margin(10, 0, 10, 0),hjust =.5),
    plot.subtitle = element_text(hjust =.5),
    axis.text=element_text(size=10),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank())+
  scale_y_continuous(labels = scales::comma)
box_total
ggsave('total_box.png',width = 12, height = 9, dpi = 'retina')

  
```
```{r}
grid_plot = ggarrange(box_mean, box_median, box_sd,box_total , 
          
          ncol = 2, nrow = 2)%>%
  ggexport(flinename =  'grid_box.png',width = 15, height = 12)

```


```{r}
ci_df = data.frame('Statistic' = as.character(),'Lower Limit' = as.numeric(), 'Upper Limit' = as.numeric())
colnames(ci_df) = c("Statistic",'Lower Limit','Upper Limit')

```

## Run bootstrapping on mean differences to find 95% confidence intervals 
```{r}
set.seed(1234)

# 100000 bootstrapping samples
N = 100000

x = numeric(N)

## Run bootstrapping samples to find difference between winning and losing teams
for (i in 1:N){
  win <- mean(sample(sal_dist$Mean_Salary[sal_dist$WinTeam == "Winning Teams"], length(w), replace = T))
  loss <- mean(sample(sal_dist$Mean_Salary[sal_dist$WinTeam == 'Losing Teams'],length(l), replace = T))
  x[i] = win-loss
}

# Develop Confidence Intervals
std = sd(x)
LL=mean(x)-1.96*std
UL=mean(x)+1.96*std
mean_ci = c('Mean Difference',LL,UL)
#cat('95% Confidence Interval: [',ci[1],',',ci[2],']')
ci_df[1,] = mean_ci

```


## Run bootstrapping on median differences to find 95% confidence intervals 
```{r}
set.seed(1234)

# 100000 bootstrapping samples
N = 100000

x = numeric(N)

## Run bootstrapping samples to find difference between winning and losing teams
for (i in 1:N){
  win <- mean(sample(sal_dist$Median.Salary[sal_dist$WinTeam == "Winning Teams"], length(w), replace = T))
  loss <- mean(sample(sal_dist$Median.Salary[sal_dist$WinTeam == 'Losing Teams'],length(l), replace = T))
  x[i] = win-loss
}

# Develop Confidence Intervals
std = sd(x)
LL=mean(x)-1.96*std
UL=mean(x)+1.96*std
median_ci = c('Mean Difference',LL,UL)
#cat('95% Confidence Interval: [',ci[1],',',ci[2],']')
ci_df[2,] = median_ci


```

```{r}
set.seed(1234)

# 100000 bootstrapping samples
N = 100000

x = numeric(N)

## Run bootstrapping samples to find difference between winning and losing teams
for (i in 1:N){
  win <- mean(sample(sal_dist$Standard_deviation[sal_dist$WinTeam == 'Winning Teams'], length(w), replace = T))
  loss <- mean(sample(sal_dist$Standard_deviation[sal_dist$WinTeam == 'Losing Teams'],length(l), replace = T))
  x[i] = win-loss
}

# Develop Confidence Intervals
std = sd(x)
LL=mean(x)-1.96*std
UL=mean(x)+1.96*std

sd_ci = c('Median Difference',LL,UL)
ci_df[3,] = sd_ci
#cat('95% Confidence Interval: [',ci[1],',',ci[2],']')



```

```{r}
set.seed(1234)

# 100000 bootstrapping samples
N = 100000

x = numeric(N)

## Run bootstrapping samples to find difference between winning and losing teams
for (i in 1:N){
  win <- mean(sample(sal_dist$Total[sal_dist$WinTeam == 'Winning Teams'], length(w), replace = T))
  loss <- mean(sample(sal_dist$Total[sal_dist$WinTeam == 'Losing Teams'],length(l), replace = T))
  x[i] = win-loss
}

# Develop Confidence Intervals
std = sd(x)
LL=mean(x)-1.96*std
UL=mean(x)+1.96*std
ci = c(LL,UL)
cat('95% Confidence Interval: [',ci[1],',',ci[2],']')

tot_ci = c('Total Difference',LL,UL)
ci_df[4,] = tot_ci

```
```{r}
ci_df$`Lower Limit` = round(as.numeric(ci_df$`Lower Limit`),-4)
ci_df$`Upper Limit` = round(as.numeric(ci_df$`Upper Limit`),-4)

ci_tb = gt(ci_df)
ci_tb = ci_tb %>%
  
# Create title & Subtitle
  tab_header(
    title = 'Confidence Intervals for Each Distribution Statistic',
    subtitle = 'Difference in Statistic of Winning and Losing Teams')%>%
  fmt_currency(
    columns = `Lower Limit`,
    currency = "USD") %>%
  fmt_currency(
    columns = `Upper Limit`,
    currency = "USD") 
ci_tb
#gtsave(acc_tb,'OX_acc_tab.html')


```
## Melt Winning dataframe in order to plot each individual team as its own line 
```{r}

# Create new data frame
df_w = data.frame(x=1:21)

# Copy salaries as new variable
sals = sal_dist$salaries[sal_dist$WinTeam == 'Winning Teams']

# Get salaries from each team and make it a new column in df
for (i in 1:length(w)){

  a = replicate(21,0)
  b = sort(sals[[i]],decreasing =T)
  a[1:length(b)] = b

  nam = paste("col", i, sep = "")
  df_w = cbind(df_w,a)
  colnames(df_w)[ncol(df_w)] = nam
  
  
}

# melt dataframe to make x's repeat & y's be salaries of each team 
df_w = melt(df_w, id = "x")
```

## Melt Losing dataframe in order to plot each individual team as its own line 
```{r}
# Create new data frame
df_l = data.frame(x=1:21)

# Copy salaries as new variable
sals = sal_dist$salaries[sal_dist$WinTeam == 'Losing Teams']

# Get salaries from each team and make it a new column in df
for (i in 1:length(l)){

  a = replicate(21,0)
  b = sort(sals[[i]],decreasing = T)
  a[1:length(b)] = b
  nam = paste("col", i, sep = "")
  
  df_l = cbind(df_l,a)
  colnames(df_l)[ncol(df_l)] = nam
  
  
}

# melt dataframe to make x's repeat & y's be salaries of each team 
df_l = melt(df_l, id = "x")

```

## Find mins and maxes for plotting ribbon
```{r}
win_max = c()
win_min = c()
lose_max = c()
lose_min = c()
for (i in 1:21){
  d = df_w$value[df_w$x == i]
  e = df_l$value[df_l$x == i]
  win_max = append(win_max,max(d))
  win_min = append(win_min,min(d))
  lose_max = append(lose_max,max(e))
  lose_min = append(lose_min,min(e))
}

```

## Plot each teams salary distribution from dataset. Separating winning and losing teams
```{r}
colors <- c("Winning Teams" = logocolors[2], "Losing Teams" = logocolors[1])

plt <- ggplot() +
  geom_line(data = df_w,aes(group = variable, x = x,y = value/100,color = "Winning Teams"),alpha = .5)+
  geom_line(data = df_l,aes(group = variable, x = x,y = value/100,color = "Losing Teams"),alpha = .5)+
  geom_ribbon(aes(x = 1:21,ymin = win_min/100, ymax = win_max/100), fill = logocolors[2],alpha = .45 )+
  geom_ribbon(aes(x = 1:21,ymin = lose_min/100, ymax = lose_max/100), fill = logocolors[1],alpha = .45)+
  labs(x = "Players",
       y = "Salary",
       color = "Legend",
       title = "Salary Distribution of NBA Teams",
       subtitle = "2009-2020 Seasons",
       caption = "Winning & losing teams split at 50% win loss record"
         ) +
  scale_color_manual(values = colors)+
  theme(plot.title = element_text(size=17, face="bold", 
    margin = margin(10, 0, 10, 0),hjust =.5),
    plot.subtitle = element_text(hjust =.5),
    plot.caption = element_text(hjust=0.5))+
  scale_y_continuous(labels = scales::comma)
plt
ggsave('Team Salary Distributions.png',plot = plt)
```


## EDA for presentation
```{r}
# Salaries Example
ex = sal[sal$TEAM == 'BKN' & sal$year == '20_21',]
ex = ex[1:13,]

plt <- ggplot(data = ex,aes(x = fct_reorder(player,-salary), y = salary))+
  geom_bar(fill = 'black',color = 'grey', stat = "identity", alpha = 0.9)+
  labs(x = "Players",
       y = "Salary",
       title = "Salary Distribution of the Brooklyn Nets",
       subtitle = "2020-2021 Season"
         ) +
  theme(plot.title = element_text(size=17, face="bold", 
    margin = margin(10, 0, 10, 0),hjust =.5),
    plot.subtitle = element_text(hjust =.5),
    axis.text.x = element_text(angle = 60, vjust = 0.99, hjust=1),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.background = element_blank())+
  scale_y_continuous(labels = scales::comma)

plt
ggsave('Nets_20_21.png',plot = plt,width = 12, height = 9, dpi = 'retina')
```


```{r}
# Salary Cap Example
cap = sal_dist[sal_dist$Season == '20_21',]
plt <- ggplot(data = cap,aes(x = fct_reorder(Team,-Total), y = Total))+
  geom_bar(fill = 'blue', stat = "identity", alpha = 0.9)+
  geom_hline(yintercept=112414000, linetype="dashed", color = "red",size=1.3)+
  geom_text(aes(28, 112414000, label = 'Salary Cap', vjust = - 1))+
  labs(x = "Teams",
       y = "Total Salary Spent",
       title = "Total Salary Spent by NBA Teams",
       subtitle = "2020-2021 Season"
         ) +
  theme(plot.title = element_text(size=17, face="bold", 
    margin = margin(10, 0, 10, 0),hjust =.5),
    plot.subtitle = element_text(hjust =.5),
    axis.text.x = element_text(angle = 60, vjust = 0.99, hjust=1),
    axis.text=element_text(size=12),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank())+
  scale_y_continuous(labels = scales::comma)

plt
ggsave('sal_cap_20_21.png',plot = plt,width = 12, height = 9, dpi = 'retina')
```

```{r}
# Data table
dt = cap[1:7,c(1,2,11,10)]
colnames(dt)[3] = 'Win Percentage'
colnames(dt)[4] = 'Salaries'

dt %>%
  kbl() %>%
  kable_material(c("striped", "hover"))

```


```{r}
# Histogram  of player salaries on winning and losing teams
colors <- c("Winning Teams" = logocolors[2], "Losing Teams" = logocolors[1])

plt <- ggplot() +
  geom_histogram(data = df_w[df_w$value>20000000,],aes(x = value,color = "Winning Teams"), fill = logocolors[2], bins = 15,alpha = .5)+
  geom_histogram(data = df_l[df_l$value>20000000,],aes(x = value,color = "Losing Teams"), fill = logocolors[1], bins = 15, alpha = .5)+
  labs(x = "Salary",
       y = "Players",
       color = "Legend",
       title = "Salaries of NBA Players > $20M",
       subtitle = "2009-2020 Seasons"
         ) +
  scale_color_manual(values = colors)+
  theme(plot.title = element_text(size=17, face="bold", 
    margin = margin(10, 0, 10, 0),hjust =.5),
    plot.subtitle = element_text(hjust =.5))+
  scale_x_continuous(labels = scales::comma)+
  scale_color_manual(values = colors)
plt
ggsave('player_sals.png',plot = plt)


```